### 2024-07-08 text/template.qmd

---
title: "Some title"
author: "Some author"
date: last-modified
format: 
  html: 
    extensions: [andrewheiss/wordcount]
references:
- id: Lovelace1842
  author:
    - family: Lovelace
      given: Ada Augusta
  citation-key: Lovelace1842
  container-title: Taylor's Scientific Memoirs
  issued:
    - year: 1842
  language: en-GB
  page: 666–731
  title: >-
    Sketch of the analytical engine invented by Charles Babbage, by LF Menabrea,
    officer of the military engineers, with notes upon the memoir by the
    translator
  type: article-journal
  volume: 3

- id: Turing1936
  author:
    - family: Turing
      given: Alan Mathison
  citation-key: Turing1936
  container-title: Journal of Math
  issue: 345–363
  issued:
    - year: 1936
  page: 230–265
  title: On Computable Numbers, with an Application to the Entscheidungsproblem
  type: article-journal
  volume: 58
---

## Introduction

Lorem ipsum dolor sit amet, consectetur adipisicing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua [@Lovelace1842]. Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia deserunt mollit anim id est laborum [@Turing1936].

| Column A | Column B |
|----------|----------|
| 7        | Apple    |
| 19       | Banana   |

: Here's a table with a caption that doesn't get counted

::: {.no-count}

### No counting

You can use the `{.no-count}` class on a [fenced div](https://quarto.org/docs/authoring/markdown-basics.html#sec-divs-and-spans) to skip counting its contents, like this section.

:::

### Shortcodes

Word counts can be directly included in your document with shortcodes, like so:[^1]

- Use `{{{< words-total >}}}` to include a count of all words: **{{< words-total >}} words**

- Use `{{{< words-body >}}}` to include a count of the words in the text body only, omitting the references, notes, and appendix: **{{< words-body >}} words**

- Use `{{{< words-ref >}}}` to include a count of the words in the reference section: **{{< words-ref >}} words**

- Use `{{{< words-append >}}}` to include a count of the words in the appendix, which must be wrapped in a div with the `#appendix-count` id (see below for an example): **{{< words-append >}} words**

- Use `{{{< words-note >}}}` to include a count of the words in the notes: **{{< words-note >}} words**. You can also reliably count words in the footnotes which would have previously been included in whichever section they were.[^2]

- Use `{{{< words-sum ARG >}}}` where `ARG` is some concatenation of the four countable areas: `body`, `ref`, `append`, and `note`. For example, `{{{< words-sum body-note >}}}` includes a count of the words in the body and notes (**{{< words-sum body-note >}} words**); `{{{< words-sum ref-append >}}}` includes a count of the words in the references and appendix (**{{< words-sum ref-append >}} words**)

[^1]: **NOTE** shortcodes `{{{< words-body >}}}`, `{{{< words-ref >}}}`, `{{{< words-append >}}}`, and `{{{< words-note >}}}` do not overlap their counts.

[^2]: Count footnote words with `{{{< words-note >}}}`: {{< words-note >}}

### Minor issues with shortcode counting

This filter may not always produce accurate counts for documents that use shortcodes, specifically shortcodes that run after this filter. If a shortcode has multiple arguments but only returns one word, then it is possible that the word count is inflated,[^3] and conversely deflated if it returns more words.

[^3]: An example in this extension is `{{{< words-sum ARG >}}}`, where `ARG` is formatted as "ref-note" or "body-append", etc., to count specific sections together. Since `words-sum` has two words to construct the shortcode but only one word is returned, the region that uses this shortcode is inflated by one.

## Conclusion

Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat. Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur.

## References [^4]

[^4]: Count reference words with `{{{< words-ref >}}}`: {{< words-ref >}}

::: {#refs}
:::

::: {#appendix-count}
## Appendix [^5] {.appendix}

Lorem ipsum dolor sit amet, consectetur adipisicing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua.
:::

[^5]: Count appendix words with `{{{< words-append >}}}`: {{< words-append >}}