### 2024-09-11 results/tables/socal_table.tex
worked on tables

\begin{table}[!h]
\centering
\begin{tabular}[t]{l|>{}l||r|r|r|r|r|r|r|r|r|r}
\hline
\multicolumn{2}{c|}{ } & \multicolumn{5}{c|}{2000} & \multicolumn{5}{c}{2020} \\
\cline{3-7} \cline{8-12}
\textbf{Variable} & \textbf{EPS Name} & \textbf{Mean} & \textbf{SD} & \textbf{P25} & \textbf{P50} & \textbf{P75} & \textbf{Mean} & \textbf{SD} & \textbf{P25} & \textbf{P50} & \textbf{P75}\\
\hline
\cellcolor{gray!6}{\% White, non-Hispanic} & \cellcolor{gray!6}{Anaheim} & \cellcolor{gray!6}{40.4} & \cellcolor{gray!6}{21.0} & \cellcolor{gray!6}{23.8} & \cellcolor{gray!6}{36.8} & \cellcolor{gray!6}{56.2} & \cellcolor{gray!6}{26.1} & \cellcolor{gray!6}{17.8} & \cellcolor{gray!6}{12.6} & \cellcolor{gray!6}{20.4} & \cellcolor{gray!6}{36.9}\\
\hline
\hline
\% White, non-Hispanic & Covina \& West Covina & 32.8 & 23.0 & 11.9 & 27.9 & 49.3 & 18.8 & 16.8 & 5.4 & 12.5 & 27.1\\
\hline
\cellcolor{gray!6}{\% White, non-Hispanic} & \cellcolor{gray!6}{East Los Angeles} & \cellcolor{gray!6}{10.3} & \cellcolor{gray!6}{11.5} & \cellcolor{gray!6}{2.9} & \cellcolor{gray!6}{6.2} & \cellcolor{gray!6}{11.8} & \cellcolor{gray!6}{6.5} & \cellcolor{gray!6}{7.1} & \cellcolor{gray!6}{2.0} & \cellcolor{gray!6}{4.1} & \cellcolor{gray!6}{8.4}\\
\hline
\% White, non-Hispanic & Glendale and Pasadena & 39.9 & 24.0 & 16.0 & 40.4 & 61.3 & 35.3 & 24.0 & 12.8 & 34.1 & 57.2\\
\hline
\cellcolor{gray!6}{\% White, non-Hispanic} & \cellcolor{gray!6}{Hollywood \& Wilshire} & \cellcolor{gray!6}{26.5} & \cellcolor{gray!6}{28.5} & \cellcolor{gray!6}{4.0} & \cellcolor{gray!6}{11.4} & \cellcolor{gray!6}{40.8} & \cellcolor{gray!6}{31.2} & \cellcolor{gray!6}{25.7} & \cellcolor{gray!6}{8.1} & \cellcolor{gray!6}{22.6} & \cellcolor{gray!6}{53.5}\\
\hline
\% White, non-Hispanic & Long Beach & 35.1 & 28.9 & 10.0 & 21.8 & 64.0 & 25.8 & 23.0 & 6.6 & 16.4 & 45.5\\
\hline
\cellcolor{gray!6}{\% White, non-Hispanic} & \cellcolor{gray!6}{Riverside, San Bernardino, \& Ontario} & \cellcolor{gray!6}{45.2} & \cellcolor{gray!6}{21.8} & \cellcolor{gray!6}{28.1} & \cellcolor{gray!6}{43.6} & \cellcolor{gray!6}{63.7} & \cellcolor{gray!6}{27.7} & \cellcolor{gray!6}{18.8} & \cellcolor{gray!6}{12.5} & \cellcolor{gray!6}{23.5} & \cellcolor{gray!6}{40.7}\\
\hline
\% White, non-Hispanic & San Fernando Valley (East) & 45.1 & 28.1 & 21.4 & 40.8 & 74.2 & 43.3 & 25.7 & 22.5 & 39.6 & 67.0\\
\hline
\cellcolor{gray!6}{\% White, non-Hispanic} & \cellcolor{gray!6}{San Fernando Valley (West)} & \cellcolor{gray!6}{58.3} & \cellcolor{gray!6}{25.6} & \cellcolor{gray!6}{40.4} & \cellcolor{gray!6}{67.0} & \cellcolor{gray!6}{78.8} & \cellcolor{gray!6}{44.0} & \cellcolor{gray!6}{24.2} & \cellcolor{gray!6}{22.8} & \cellcolor{gray!6}{46.6} & \cellcolor{gray!6}{64.7}\\
\hline
\% White, non-Hispanic & Santa Ana & 52.4 & 29.0 & 24.5 & 58.3 & 77.9 & 42.0 & 26.3 & 17.0 & 44.1 & 64.6\\
\hline
\cellcolor{gray!6}{\% White, non-Hispanic} & \cellcolor{gray!6}{South \& South Central Los Angeles} & \cellcolor{gray!6}{7.3} & \cellcolor{gray!6}{13.6} & \cellcolor{gray!6}{1.1} & \cellcolor{gray!6}{2.3} & \cellcolor{gray!6}{5.5} & \cellcolor{gray!6}{6.3} & \cellcolor{gray!6}{11.4} & \cellcolor{gray!6}{0.6} & \cellcolor{gray!6}{1.9} & \cellcolor{gray!6}{6.6}\\
\hline
\% White, non-Hispanic & South Bay & 41.3 & 29.4 & 12.4 & 41.3 & 68.3 & 29.1 & 24.5 & 7.1 & 22.1 & 51.4\\
\hline
\cellcolor{gray!6}{\% White, non-Hispanic} & \cellcolor{gray!6}{South Orange Co} & \cellcolor{gray!6}{66.0} & \cellcolor{gray!6}{18.0} & \cellcolor{gray!6}{57.0} & \cellcolor{gray!6}{70.4} & \cellcolor{gray!6}{78.1} & \cellcolor{gray!6}{52.9} & \cellcolor{gray!6}{18.1} & \cellcolor{gray!6}{40.7} & \cellcolor{gray!6}{55.0} & \cellcolor{gray!6}{66.3}\\
\hline
\% White, non-Hispanic & West Los Angeles \& West Beach & 63.9 & 24.4 & 43.0 & 72.9 & 84.0 & 60.2 & 20.2 & 45.9 & 62.9 & 75.5\\
\hline
\cellcolor{gray!6}{\% White, non-Hispanic} & \cellcolor{gray!6}{Whittier \& North Orange Co} & \cellcolor{gray!6}{35.4} & \cellcolor{gray!6}{20.5} & \cellcolor{gray!6}{18.5} & \cellcolor{gray!6}{31.2} & \cellcolor{gray!6}{50.7} & \cellcolor{gray!6}{22.4} & \cellcolor{gray!6}{15.3} & \cellcolor{gray!6}{10.9} & \cellcolor{gray!6}{17.5} & \cellcolor{gray!6}{32.7}\\
\hline
\% Black, non-Hispanic & Anaheim & 3.9 & 5.3 & 1.3 & 2.3 & 4.4 & 4.4 & 5.6 & 0.8 & 2.3 & 6.0\\
\hline
\hline
\cellcolor{gray!6}{\% Black, non-Hispanic} & \cellcolor{gray!6}{Covina \& West Covina} & \cellcolor{gray!6}{4.1} & \cellcolor{gray!6}{4.2} & \cellcolor{gray!6}{1.4} & \cellcolor{gray!6}{3.0} & \cellcolor{gray!6}{5.2} & \cellcolor{gray!6}{3.2} & \cellcolor{gray!6}{3.1} & \cellcolor{gray!6}{0.8} & \cellcolor{gray!6}{2.3} & \cellcolor{gray!6}{4.7}\\
\hline
\% Black, non-Hispanic & East Los Angeles & 1.1 & 2.8 & 0.2 & 0.5 & 1.0 & 1.3 & 2.8 & 0.0 & 0.4 & 1.4\\
\hline
\cellcolor{gray!6}{\% Black, non-Hispanic} & \cellcolor{gray!6}{Glendale and Pasadena} & \cellcolor{gray!6}{4.4} & \cellcolor{gray!6}{8.5} & \cellcolor{gray!6}{0.8} & \cellcolor{gray!6}{1.7} & \cellcolor{gray!6}{3.2} & \cellcolor{gray!6}{3.4} & \cellcolor{gray!6}{5.2} & \cellcolor{gray!6}{0.5} & \cellcolor{gray!6}{1.8} & \cellcolor{gray!6}{4.1}\\
\hline
\% Black, non-Hispanic & Hollywood \& Wilshire & 13.0 & 21.0 & 2.0 & 3.3 & 8.5 & 11.1 & 16.8 & 1.9 & 4.5 & 10.5\\
\hline
\cellcolor{gray!6}{\% Black, non-Hispanic} & \cellcolor{gray!6}{Long Beach} & \cellcolor{gray!6}{13.1} & \cellcolor{gray!6}{14.4} & \cellcolor{gray!6}{3.0} & \cellcolor{gray!6}{8.9} & \cellcolor{gray!6}{18.8} & \cellcolor{gray!6}{11.4} & \cellcolor{gray!6}{13.0} & \cellcolor{gray!6}{3.6} & \cellcolor{gray!6}{8.1} & \cellcolor{gray!6}{15.1}\\
\hline
\% Black, non-Hispanic & Riverside, San Bernardino, \& Ontario & 8.2 & 7.0 & 3.2 & 5.8 & 11.4 & 7.1 & 6.4 & 2.4 & 5.4 & 9.8\\
\hline
\cellcolor{gray!6}{\% Black, non-Hispanic} & \cellcolor{gray!6}{San Fernando Valley (East)} & \cellcolor{gray!6}{4.0} & \cellcolor{gray!6}{2.3} & \cellcolor{gray!6}{2.3} & \cellcolor{gray!6}{3.7} & \cellcolor{gray!6}{5.3} & \cellcolor{gray!6}{4.3} & \cellcolor{gray!6}{3.8} & \cellcolor{gray!6}{1.6} & \cellcolor{gray!6}{3.3} & \cellcolor{gray!6}{5.7}\\
\hline
\% Black, non-Hispanic & San Fernando Valley (West) & 3.2 & 3.1 & 1.2 & 2.5 & 3.8 & 3.8 & 3.9 & 1.1 & 2.7 & 5.4\\
\hline
\cellcolor{gray!6}{\% Black, non-Hispanic} & \cellcolor{gray!6}{Santa Ana} & \cellcolor{gray!6}{1.2} & \cellcolor{gray!6}{1.8} & \cellcolor{gray!6}{0.6} & \cellcolor{gray!6}{0.9} & \cellcolor{gray!6}{1.4} & \cellcolor{gray!6}{1.3} & \cellcolor{gray!6}{1.9} & \cellcolor{gray!6}{0.1} & \cellcolor{gray!6}{0.7} & \cellcolor{gray!6}{1.6}\\
\hline
\% Black, non-Hispanic & South \& South Central Los Angeles & 24.7 & 24.4 & 2.6 & 17.7 & 39.6 & 17.5 & 18.8 & 2.7 & 11.6 & 25.3\\
\hline
\cellcolor{gray!6}{\% Black, non-Hispanic} & \cellcolor{gray!6}{South Bay} & \cellcolor{gray!6}{13.2} & \cellcolor{gray!6}{20.2} & \cellcolor{gray!6}{1.5} & \cellcolor{gray!6}{4.1} & \cellcolor{gray!6}{15.2} & \cellcolor{gray!6}{12.1} & \cellcolor{gray!6}{16.9} & \cellcolor{gray!6}{1.7} & \cellcolor{gray!6}{5.7} & \cellcolor{gray!6}{14.3}\\
\hline
\% Black, non-Hispanic & South Orange Co & 1.4 & 1.2 & 0.7 & 1.2 & 1.8 & 1.5 & 1.8 & 0.3 & 1.0 & 2.3\\
\hline
\cellcolor{gray!6}{\% Black, non-Hispanic} & \cellcolor{gray!6}{West Los Angeles \& West Beach} & \cellcolor{gray!6}{6.3} & \cellcolor{gray!6}{11.7} & \cellcolor{gray!6}{1.4} & \cellcolor{gray!6}{2.4} & \cellcolor{gray!6}{4.7} & \cellcolor{gray!6}{5.4} & \cellcolor{gray!6}{8.1} & \cellcolor{gray!6}{1.1} & \cellcolor{gray!6}{2.9} & \cellcolor{gray!6}{6.2}\\
\hline
\% Black, non-Hispanic & Whittier \& North Orange Co & 2.2 & 2.3 & 0.9 & 1.5 & 2.5 & 2.4 & 2.7 & 0.4 & 1.5 & 3.0\\
\hline
\cellcolor{gray!6}{\% Hispanic} & \cellcolor{gray!6}{Anaheim} & \cellcolor{gray!6}{34.0} & \cellcolor{gray!6}{22.0} & \cellcolor{gray!6}{16.3} & \cellcolor{gray!6}{29.7} & \cellcolor{gray!6}{49.1} & \cellcolor{gray!6}{41.3} & \cellcolor{gray!6}{22.7} & \cellcolor{gray!6}{23.1} & \cellcolor{gray!6}{38.6} & \cellcolor{gray!6}{58.1}\\
\hline
\hline
\% Hispanic & Covina \& West Covina & 43.6 & 25.6 & 20.6 & 39.6 & 68.7 & 51.7 & 23.5 & 30.8 & 54.3 & 71.3\\
\hline
\cellcolor{gray!6}{\% Hispanic} & \cellcolor{gray!6}{East Los Angeles} & \cellcolor{gray!6}{66.6} & \cellcolor{gray!6}{27.9} & \cellcolor{gray!6}{41.6} & \cellcolor{gray!6}{77.9} & \cellcolor{gray!6}{91.0} & \cellcolor{gray!6}{64.4} & \cellcolor{gray!6}{28.9} & \cellcolor{gray!6}{36.7} & \cellcolor{gray!6}{73.2} & \cellcolor{gray!6}{91.5}\\
\hline
\% Hispanic & Glendale and Pasadena & 34.4 & 25.8 & 13.8 & 26.3 & 53.5 & 36.8 & 26.9 & 16.3 & 26.3 & 53.4\\
\hline
\cellcolor{gray!6}{\% Hispanic} & \cellcolor{gray!6}{Hollywood \& Wilshire} & \cellcolor{gray!6}{43.7} & \cellcolor{gray!6}{28.3} & \cellcolor{gray!6}{15.8} & \cellcolor{gray!6}{44.4} & \cellcolor{gray!6}{68.5} & \cellcolor{gray!6}{38.2} & \cellcolor{gray!6}{24.6} & \cellcolor{gray!6}{15.7} & \cellcolor{gray!6}{34.7} & \cellcolor{gray!6}{57.1}\\
\hline
\% Hispanic & Long Beach & 37.1 & 24.3 & 16.3 & 32.1 & 53.1 & 45.8 & 23.3 & 26.6 & 43.5 & 63.4\\
\hline
\cellcolor{gray!6}{\% Hispanic} & \cellcolor{gray!6}{Riverside, San Bernardino, \& Ontario} & \cellcolor{gray!6}{38.3} & \cellcolor{gray!6}{20.5} & \cellcolor{gray!6}{21.7} & \cellcolor{gray!6}{34.2} & \cellcolor{gray!6}{52.9} & \cellcolor{gray!6}{53.8} & \cellcolor{gray!6}{22.0} & \cellcolor{gray!6}{36.3} & \cellcolor{gray!6}{54.1} & \cellcolor{gray!6}{72.4}\\
\hline
\% Hispanic & San Fernando Valley (East) & 40.2 & 27.5 & 11.3 & 35.8 & 62.5 & 40.6 & 27.2 & 14.8 & 37.3 & 60.9\\
\hline
\cellcolor{gray!6}{\% Hispanic} & \cellcolor{gray!6}{San Fernando Valley (West)} & \cellcolor{gray!6}{27.4} & \cellcolor{gray!6}{25.0} & \cellcolor{gray!6}{9.4} & \cellcolor{gray!6}{16.4} & \cellcolor{gray!6}{38.8} & \cellcolor{gray!6}{36.8} & \cellcolor{gray!6}{25.9} & \cellcolor{gray!6}{16.2} & \cellcolor{gray!6}{28.8} & \cellcolor{gray!6}{56.9}\\
\hline
\% Hispanic & Santa Ana & 29.6 & 29.2 & 8.1 & 15.0 & 43.0 & 33.0 & 27.9 & 12.3 & 21.2 & 46.6\\
\hline
\cellcolor{gray!6}{\% Hispanic} & \cellcolor{gray!6}{South \& South Central Los Angeles} & \cellcolor{gray!6}{63.6} & \cellcolor{gray!6}{26.9} & \cellcolor{gray!6}{46.5} & \cellcolor{gray!6}{65.9} & \cellcolor{gray!6}{88.3} & \cellcolor{gray!6}{70.5} & \cellcolor{gray!6}{24.8} & \cellcolor{gray!6}{60.2} & \cellcolor{gray!6}{75.7} & \cellcolor{gray!6}{90.2}\\
\hline
\% Hispanic & South Bay & 25.6 & 20.6 & 8.8 & 19.8 & 39.0 & 33.2 & 22.5 & 12.9 & 29.8 & 51.8\\
\hline
\cellcolor{gray!6}{\% Hispanic} & \cellcolor{gray!6}{South Orange Co} & \cellcolor{gray!6}{18.0} & \cellcolor{gray!6}{16.2} & \cellcolor{gray!6}{7.6} & \cellcolor{gray!6}{11.6} & \cellcolor{gray!6}{21.3} & \cellcolor{gray!6}{22.2} & \cellcolor{gray!6}{17.0} & \cellcolor{gray!6}{10.4} & \cellcolor{gray!6}{16.8} & \cellcolor{gray!6}{27.2}\\
\hline
\% Hispanic & West Los Angeles \& West Beach & 16.2 & 16.6 & 4.6 & 9.4 & 21.1 & 15.8 & 13.4 & 6.3 & 12.2 & 20.4\\
\hline
\cellcolor{gray!6}{\% Hispanic} & \cellcolor{gray!6}{Whittier \& North Orange Co} & \cellcolor{gray!6}{47.5} & \cellcolor{gray!6}{24.9} & \cellcolor{gray!6}{27.2} & \cellcolor{gray!6}{46.7} & \cellcolor{gray!6}{67.1} & \cellcolor{gray!6}{56.0} & \cellcolor{gray!6}{23.6} & \cellcolor{gray!6}{39.3} & \cellcolor{gray!6}{59.2} & \cellcolor{gray!6}{75.7}\\
\hline
\% API, non-Hispanic & Anaheim & 18.8 & 15.7 & 8.1 & 14.3 & 23.2 & 24.5 & 18.1 & 11.2 & 19.4 & 34.2\\
\hline
\hline
\cellcolor{gray!6}{\% API, non-Hispanic} & \cellcolor{gray!6}{Covina \& West Covina} & \cellcolor{gray!6}{17.1} & \cellcolor{gray!6}{17.1} & \cellcolor{gray!6}{5.8} & \cellcolor{gray!6}{9.7} & \cellcolor{gray!6}{21.9} & \cellcolor{gray!6}{23.7} & \cellcolor{gray!6}{19.3} & \cellcolor{gray!6}{10.0} & \cellcolor{gray!6}{16.7} & \cellcolor{gray!6}{31.9}\\
\hline
\% API, non-Hispanic & East Los Angeles & 20.5 & 21.5 & 1.8 & 11.2 & 37.9 & 26.5 & 26.0 & 2.7 & 16.4 & 47.9\\
\hline
\cellcolor{gray!6}{\% API, non-Hispanic} & \cellcolor{gray!6}{Glendale and Pasadena} & \cellcolor{gray!6}{17.4} & \cellcolor{gray!6}{15.5} & \cellcolor{gray!6}{6.2} & \cellcolor{gray!6}{11.6} & \cellcolor{gray!6}{23.1} & \cellcolor{gray!6}{20.9} & \cellcolor{gray!6}{19.8} & \cellcolor{gray!6}{7.1} & \cellcolor{gray!6}{13.0} & \cellcolor{gray!6}{27.3}\\
\hline
\% API, non-Hispanic & Hollywood \& Wilshire & 13.8 & 12.9 & 4.6 & 9.5 & 18.9 & 15.7 & 14.0 & 5.7 & 11.4 & 20.1\\
\hline
\cellcolor{gray!6}{\% API, non-Hispanic} & \cellcolor{gray!6}{Long Beach} & \cellcolor{gray!6}{11.4} & \cellcolor{gray!6}{10.1} & \cellcolor{gray!6}{4.7} & \cellcolor{gray!6}{8.1} & \cellcolor{gray!6}{15.3} & \cellcolor{gray!6}{13.2} & \cellcolor{gray!6}{11.3} & \cellcolor{gray!6}{4.9} & \cellcolor{gray!6}{10.2} & \cellcolor{gray!6}{18.2}\\
\hline
\% API, non-Hispanic & Riverside, San Bernardino, \& Ontario & 5.2 & 6.5 & 1.8 & 3.5 & 6.0 & 8.2 & 10.0 & 2.0 & 5.0 & 10.4\\
\hline
\cellcolor{gray!6}{\% API, non-Hispanic} & \cellcolor{gray!6}{San Fernando Valley (East)} & \cellcolor{gray!6}{7.2} & \cellcolor{gray!6}{4.9} & \cellcolor{gray!6}{4.6} & \cellcolor{gray!6}{5.7} & \cellcolor{gray!6}{7.7} & \cellcolor{gray!6}{8.3} & \cellcolor{gray!6}{5.4} & \cellcolor{gray!6}{4.7} & \cellcolor{gray!6}{6.9} & \cellcolor{gray!6}{10.2}\\
\hline
\% API, non-Hispanic & San Fernando Valley (West) & 8.0 & 5.9 & 4.2 & 6.5 & 10.8 & 11.6 & 8.1 & 5.3 & 10.4 & 16.3\\
\hline
\cellcolor{gray!6}{\% API, non-Hispanic} & \cellcolor{gray!6}{Santa Ana} & \cellcolor{gray!6}{14.1} & \cellcolor{gray!6}{13.6} & \cellcolor{gray!6}{4.5} & \cellcolor{gray!6}{9.2} & \cellcolor{gray!6}{18.1} & \cellcolor{gray!6}{20.1} & \cellcolor{gray!6}{18.3} & \cellcolor{gray!6}{6.5} & \cellcolor{gray!6}{14.3} & \cellcolor{gray!6}{26.5}\\
\hline
\% API, non-Hispanic & South \& South Central Los Angeles & 2.9 & 5.3 & 0.3 & 0.8 & 2.7 & 3.9 & 6.9 & 0.2 & 1.2 & 4.4\\
\hline
\cellcolor{gray!6}{\% API, non-Hispanic} & \cellcolor{gray!6}{South Bay} & \cellcolor{gray!6}{16.6} & \cellcolor{gray!6}{13.7} & \cellcolor{gray!6}{6.1} & \cellcolor{gray!6}{11.0} & \cellcolor{gray!6}{26.8} & \cellcolor{gray!6}{21.2} & \cellcolor{gray!6}{15.9} & \cellcolor{gray!6}{8.2} & \cellcolor{gray!6}{17.8} & \cellcolor{gray!6}{30.3}\\
\hline
\% API, non-Hispanic & South Orange Co & 11.1 & 9.8 & 5.3 & 8.5 & 13.6 & 18.8 & 15.1 & 9.0 & 14.1 & 23.6\\
\hline
\cellcolor{gray!6}{\% API, non-Hispanic} & \cellcolor{gray!6}{West Los Angeles \& West Beach} & \cellcolor{gray!6}{9.9} & \cellcolor{gray!6}{7.9} & \cellcolor{gray!6}{4.9} & \cellcolor{gray!6}{7.5} & \cellcolor{gray!6}{11.7} & \cellcolor{gray!6}{13.1} & \cellcolor{gray!6}{9.5} & \cellcolor{gray!6}{6.5} & \cellcolor{gray!6}{9.8} & \cellcolor{gray!6}{17.6}\\
\hline
\% API, non-Hispanic & Whittier \& North Orange Co & 12.6 & 15.4 & 3.6 & 8.2 & 12.9 & 15.9 & 16.2 & 5.2 & 12.0 & 18.0\\
\hline
\cellcolor{gray!6}{\% Native, non-Hispanic} & \cellcolor{gray!6}{Anaheim} & \cellcolor{gray!6}{0.3} & \cellcolor{gray!6}{0.2} & \cellcolor{gray!6}{0.2} & \cellcolor{gray!6}{0.3} & \cellcolor{gray!6}{0.4} & \cellcolor{gray!6}{0.5} & \cellcolor{gray!6}{6.0} & \cellcolor{gray!6}{0.0} & \cellcolor{gray!6}{0.0} & \cellcolor{gray!6}{0.2}\\
\hline
\hline
\% Native, non-Hispanic & Covina \& West Covina & 0.3 & 0.2 & 0.2 & 0.3 & 0.4 & 0.2 & 0.7 & 0.0 & 0.0 & 0.2\\
\hline
\cellcolor{gray!6}{\% Native, non-Hispanic} & \cellcolor{gray!6}{East Los Angeles} & \cellcolor{gray!6}{0.2} & \cellcolor{gray!6}{0.2} & \cellcolor{gray!6}{0.1} & \cellcolor{gray!6}{0.2} & \cellcolor{gray!6}{0.3} & \cellcolor{gray!6}{0.2} & \cellcolor{gray!6}{0.6} & \cellcolor{gray!6}{0.0} & \cellcolor{gray!6}{0.0} & \cellcolor{gray!6}{0.2}\\
\hline
\% Native, non-Hispanic & Glendale and Pasadena & 0.3 & 0.2 & 0.1 & 0.2 & 0.3 & 0.2 & 0.5 & 0.0 & 0.0 & 0.1\\
\hline
\cellcolor{gray!6}{\% Native, non-Hispanic} & \cellcolor{gray!6}{Hollywood \& Wilshire} & \cellcolor{gray!6}{0.2} & \cellcolor{gray!6}{0.2} & \cellcolor{gray!6}{0.1} & \cellcolor{gray!6}{0.2} & \cellcolor{gray!6}{0.3} & \cellcolor{gray!6}{0.2} & \cellcolor{gray!6}{0.5} & \cellcolor{gray!6}{0.0} & \cellcolor{gray!6}{0.0} & \cellcolor{gray!6}{0.0}\\
\hline
\% Native, non-Hispanic & Long Beach & 0.4 & 0.3 & 0.2 & 0.3 & 0.5 & 0.6 & 6.9 & 0.0 & 0.0 & 0.1\\
\hline
\cellcolor{gray!6}{\% Native, non-Hispanic} & \cellcolor{gray!6}{Riverside, San Bernardino, \& Ontario} & \cellcolor{gray!6}{0.5} & \cellcolor{gray!6}{0.4} & \cellcolor{gray!6}{0.3} & \cellcolor{gray!6}{0.5} & \cellcolor{gray!6}{0.6} & \cellcolor{gray!6}{0.3} & \cellcolor{gray!6}{0.6} & \cellcolor{gray!6}{0.0} & \cellcolor{gray!6}{0.0} & \cellcolor{gray!6}{0.3}\\
\hline
\% Native, non-Hispanic & San Fernando Valley (East) & 0.2 & 0.1 & 0.1 & 0.2 & 0.3 & 0.1 & 0.3 & 0.0 & 0.0 & 0.0\\
\hline
\cellcolor{gray!6}{\% Native, non-Hispanic} & \cellcolor{gray!6}{San Fernando Valley (West)} & \cellcolor{gray!6}{0.3} & \cellcolor{gray!6}{0.2} & \cellcolor{gray!6}{0.2} & \cellcolor{gray!6}{0.3} & \cellcolor{gray!6}{0.4} & \cellcolor{gray!6}{0.2} & \cellcolor{gray!6}{0.5} & \cellcolor{gray!6}{0.0} & \cellcolor{gray!6}{0.0} & \cellcolor{gray!6}{0.1}\\
\hline
\% Native, non-Hispanic & Santa Ana & 0.3 & 0.2 & 0.2 & 0.3 & 0.4 & 0.2 & 0.4 & 0.0 & 0.0 & 0.2\\
\hline
\cellcolor{gray!6}{\% Native, non-Hispanic} & \cellcolor{gray!6}{South \& South Central Los Angeles} & \cellcolor{gray!6}{0.2} & \cellcolor{gray!6}{0.2} & \cellcolor{gray!6}{0.1} & \cellcolor{gray!6}{0.2} & \cellcolor{gray!6}{0.3} & \cellcolor{gray!6}{0.1} & \cellcolor{gray!6}{0.4} & \cellcolor{gray!6}{0.0} & \cellcolor{gray!6}{0.0} & \cellcolor{gray!6}{0.0}\\
\hline
\% Native, non-Hispanic & South Bay & 0.2 & 0.1 & 0.1 & 0.2 & 0.3 & 0.2 & 0.5 & 0.0 & 0.0 & 0.0\\
\hline
\cellcolor{gray!6}{\% Native, non-Hispanic} & \cellcolor{gray!6}{South Orange Co} & \cellcolor{gray!6}{0.3} & \cellcolor{gray!6}{0.2} & \cellcolor{gray!6}{0.2} & \cellcolor{gray!6}{0.2} & \cellcolor{gray!6}{0.4} & \cellcolor{gray!6}{0.2} & \cellcolor{gray!6}{0.4} & \cellcolor{gray!6}{0.0} & \cellcolor{gray!6}{0.0} & \cellcolor{gray!6}{0.0}\\
\hline
\% Native, non-Hispanic & West Los Angeles \& West Beach & 0.2 & 0.2 & 0.1 & 0.2 & 0.3 & 0.2 & 0.4 & 0.0 & 0.0 & 0.0\\
\hline
\cellcolor{gray!6}{\% Native, non-Hispanic} & \cellcolor{gray!6}{Whittier \& North Orange Co} & \cellcolor{gray!6}{0.3} & \cellcolor{gray!6}{0.2} & \cellcolor{gray!6}{0.2} & \cellcolor{gray!6}{0.3} & \cellcolor{gray!6}{0.4} & \cellcolor{gray!6}{0.8} & \cellcolor{gray!6}{7.7} & \cellcolor{gray!6}{0.0} & \cellcolor{gray!6}{0.0} & \cellcolor{gray!6}{0.3}\\
\hline
pct\_nhisp\_multi & Anaheim & 2.4 & 0.9 & 1.8 & 2.4 & 3.0 & 2.9 & 2.3 & 1.1 & 2.1 & 4.5\\
\hline
\hline
\cellcolor{gray!6}{pct\_nhisp\_multi} & \cellcolor{gray!6}{Covina \& West Covina} & \cellcolor{gray!6}{2.0} & \cellcolor{gray!6}{0.8} & \cellcolor{gray!6}{1.4} & \cellcolor{gray!6}{2.0} & \cellcolor{gray!6}{2.6} & \cellcolor{gray!6}{2.0} & \cellcolor{gray!6}{2.1} & \cellcolor{gray!6}{0.4} & \cellcolor{gray!6}{1.5} & \cellcolor{gray!6}{2.9}\\
\hline
pct\_nhisp\_multi & East Los Angeles & 1.1 & 0.9 & 0.4 & 0.9 & 1.7 & 0.9 & 1.2 & 0.0 & 0.4 & 1.4\\
\hline
\cellcolor{gray!6}{pct\_nhisp\_multi} & \cellcolor{gray!6}{Glendale and Pasadena} & \cellcolor{gray!6}{3.5} & \cellcolor{gray!6}{2.8} & \cellcolor{gray!6}{1.9} & \cellcolor{gray!6}{2.8} & \cellcolor{gray!6}{3.7} & \cellcolor{gray!6}{3.0} & \cellcolor{gray!6}{2.7} & \cellcolor{gray!6}{1.0} & \cellcolor{gray!6}{2.4} & \cellcolor{gray!6}{4.2}\\
\hline
pct\_nhisp\_multi & Hollywood \& Wilshire & 2.6 & 2.0 & 1.3 & 2.1 & 3.1 & 3.2 & 2.9 & 1.3 & 2.4 & 4.6\\
\hline
\cellcolor{gray!6}{pct\_nhisp\_multi} & \cellcolor{gray!6}{Long Beach} & \cellcolor{gray!6}{2.6} & \cellcolor{gray!6}{1.0} & \cellcolor{gray!6}{2.0} & \cellcolor{gray!6}{2.7} & \cellcolor{gray!6}{3.3} & \cellcolor{gray!6}{2.9} & \cellcolor{gray!6}{2.3} & \cellcolor{gray!6}{1.2} & \cellcolor{gray!6}{2.5} & \cellcolor{gray!6}{4.2}\\
\hline
pct\_nhisp\_multi & Riverside, San Bernardino, \& Ontario & 2.4 & 0.9 & 1.8 & 2.3 & 3.0 & 2.6 & 2.4 & 0.7 & 1.9 & 3.7\\
\hline
\cellcolor{gray!6}{pct\_nhisp\_multi} & \cellcolor{gray!6}{San Fernando Valley (East)} & \cellcolor{gray!6}{3.0} & \cellcolor{gray!6}{1.4} & \cellcolor{gray!6}{2.1} & \cellcolor{gray!6}{2.9} & \cellcolor{gray!6}{3.7} & \cellcolor{gray!6}{3.1} & \cellcolor{gray!6}{2.9} & \cellcolor{gray!6}{1.0} & \cellcolor{gray!6}{2.7} & \cellcolor{gray!6}{4.4}\\
\hline
pct\_nhisp\_multi & San Fernando Valley (West) & 2.6 & 1.5 & 1.7 & 2.4 & 3.2 & 3.1 & 2.6 & 1.1 & 2.7 & 4.1\\
\hline
\cellcolor{gray!6}{pct\_nhisp\_multi} & \cellcolor{gray!6}{Santa Ana} & \cellcolor{gray!6}{2.2} & \cellcolor{gray!6}{1.2} & \cellcolor{gray!6}{1.4} & \cellcolor{gray!6}{2.1} & \cellcolor{gray!6}{2.9} & \cellcolor{gray!6}{3.2} & \cellcolor{gray!6}{2.9} & \cellcolor{gray!6}{1.1} & \cellcolor{gray!6}{2.6} & \cellcolor{gray!6}{4.9}\\
\hline
pct\_nhisp\_multi & South \& South Central Los Angeles & 1.1 & 1.1 & 0.4 & 0.8 & 1.5 & 1.3 & 2.0 & 0.0 & 0.5 & 1.8\\
\hline
\cellcolor{gray!6}{pct\_nhisp\_multi} & \cellcolor{gray!6}{South Bay} & \cellcolor{gray!6}{2.7} & \cellcolor{gray!6}{1.0} & \cellcolor{gray!6}{2.1} & \cellcolor{gray!6}{2.8} & \cellcolor{gray!6}{3.5} & \cellcolor{gray!6}{3.8} & \cellcolor{gray!6}{2.9} & \cellcolor{gray!6}{1.5} & \cellcolor{gray!6}{3.2} & \cellcolor{gray!6}{5.6}\\
\hline
pct\_nhisp\_multi & South Orange Co & 3.0 & 6.2 & 1.8 & 2.4 & 3.2 & 4.1 & 2.7 & 2.1 & 3.7 & 5.5\\
\hline
\cellcolor{gray!6}{pct\_nhisp\_multi} & \cellcolor{gray!6}{West Los Angeles \& West Beach} & \cellcolor{gray!6}{3.2} & \cellcolor{gray!6}{1.3} & \cellcolor{gray!6}{2.4} & \cellcolor{gray!6}{3.1} & \cellcolor{gray!6}{3.9} & \cellcolor{gray!6}{4.8} & \cellcolor{gray!6}{2.9} & \cellcolor{gray!6}{2.7} & \cellcolor{gray!6}{4.2} & \cellcolor{gray!6}{6.7}\\
\hline
pct\_nhisp\_multi & Whittier \& North Orange Co & 1.8 & 0.7 & 1.3 & 1.9 & 2.3 & 1.9 & 1.7 & 0.5 & 1.5 & 2.8\\
\hline
\cellcolor{gray!6}{Median income} & \cellcolor{gray!6}{Anaheim} & \cellcolor{gray!6}{104.0} & \cellcolor{gray!6}{42.0} & \cellcolor{gray!6}{78.0} & \cellcolor{gray!6}{94.0} & \cellcolor{gray!6}{119.0} & \cellcolor{gray!6}{108.0} & \cellcolor{gray!6}{41.0} & \cellcolor{gray!6}{78.0} & \cellcolor{gray!6}{99.0} & \cellcolor{gray!6}{129.0}\\
\hline
\hline
Median income & Covina \& West Covina & 103.0 & 36.0 & 79.0 & 96.0 & 121.0 & 103.0 & 32.0 & 82.0 & 99.0 & 120.0\\
\hline
\cellcolor{gray!6}{Median income} & \cellcolor{gray!6}{East Los Angeles} & \cellcolor{gray!6}{70.0} & \cellcolor{gray!6}{26.0} & \cellcolor{gray!6}{53.0} & \cellcolor{gray!6}{64.0} & \cellcolor{gray!6}{81.0} & \cellcolor{gray!6}{76.0} & \cellcolor{gray!6}{25.0} & \cellcolor{gray!6}{60.0} & \cellcolor{gray!6}{72.0} & \cellcolor{gray!6}{87.0}\\
\hline
Median income & Glendale and Pasadena & 96.0 & 40.0 & 69.0 & 86.0 & 110.0 & 102.0 & 40.0 & 73.0 & 96.0 & 121.0\\
\hline
\cellcolor{gray!6}{Median income} & \cellcolor{gray!6}{Hollywood \& Wilshire} & \cellcolor{gray!6}{68.0} & \cellcolor{gray!6}{42.0} & \cellcolor{gray!6}{41.0} & \cellcolor{gray!6}{57.0} & \cellcolor{gray!6}{86.0} & \cellcolor{gray!6}{84.0} & \cellcolor{gray!6}{43.0} & \cellcolor{gray!6}{52.0} & \cellcolor{gray!6}{75.0} & \cellcolor{gray!6}{105.0}\\
\hline
Median income & Long Beach & 84.0 & 41.0 & 54.0 & 79.0 & 108.0 & 90.0 & 36.0 & 62.0 & 85.0 & 111.0\\
\hline
\cellcolor{gray!6}{Median income} & \cellcolor{gray!6}{Riverside, San Bernardino, \& Ontario} & \cellcolor{gray!6}{89.0} & \cellcolor{gray!6}{36.0} & \cellcolor{gray!6}{64.0} & \cellcolor{gray!6}{81.0} & \cellcolor{gray!6}{111.0} & \cellcolor{gray!6}{95.0} & \cellcolor{gray!6}{37.0} & \cellcolor{gray!6}{68.0} & \cellcolor{gray!6}{88.0} & \cellcolor{gray!6}{119.0}\\
\hline
Median income & San Fernando Valley (East) & 94.0 & 62.0 & 60.0 & 75.0 & 97.0 & 99.0 & 56.0 & 61.0 & 83.0 & 116.0\\
\hline
\cellcolor{gray!6}{Median income} & \cellcolor{gray!6}{San Fernando Valley (West)} & \cellcolor{gray!6}{124.0} & \cellcolor{gray!6}{52.0} & \cellcolor{gray!6}{87.0} & \cellcolor{gray!6}{113.0} & \cellcolor{gray!6}{156.0} & \cellcolor{gray!6}{121.0} & \cellcolor{gray!6}{46.0} & \cellcolor{gray!6}{87.0} & \cellcolor{gray!6}{116.0} & \cellcolor{gray!6}{146.0}\\
\hline
Median income & Santa Ana & 116.0 & 44.0 & 86.0 & 107.0 & 141.0 & 119.0 & 44.0 & 87.0 & 112.0 & 142.0\\
\hline
\cellcolor{gray!6}{Median income} & \cellcolor{gray!6}{South \& South Central Los Angeles} & \cellcolor{gray!6}{59.0} & \cellcolor{gray!6}{23.0} & \cellcolor{gray!6}{44.0} & \cellcolor{gray!6}{55.0} & \cellcolor{gray!6}{68.0} & \cellcolor{gray!6}{67.0} & \cellcolor{gray!6}{27.0} & \cellcolor{gray!6}{50.0} & \cellcolor{gray!6}{62.0} & \cellcolor{gray!6}{78.0}\\
\hline
Median income & South Bay & 111.0 & 52.0 & 73.0 & 99.0 & 134.0 & 115.0 & 51.0 & 78.0 & 103.0 & 141.0\\
\hline
\cellcolor{gray!6}{Median income} & \cellcolor{gray!6}{South Orange Co} & \cellcolor{gray!6}{140.0} & \cellcolor{gray!6}{49.0} & \cellcolor{gray!6}{104.0} & \cellcolor{gray!6}{137.0} & \cellcolor{gray!6}{167.0} & \cellcolor{gray!6}{142.0} & \cellcolor{gray!6}{46.0} & \cellcolor{gray!6}{109.0} & \cellcolor{gray!6}{140.0} & \cellcolor{gray!6}{171.0}\\
\hline
Median income & West Los Angeles \& West Beach & 132.0 & 78.0 & 80.0 & 104.0 & 168.0 & 137.0 & 61.0 & 98.0 & 122.0 & 165.0\\
\hline
\cellcolor{gray!6}{Median income} & \cellcolor{gray!6}{Whittier \& North Orange Co} & \cellcolor{gray!6}{101.0} & \cellcolor{gray!6}{29.0} & \cellcolor{gray!6}{80.0} & \cellcolor{gray!6}{94.0} & \cellcolor{gray!6}{116.0} & \cellcolor{gray!6}{105.0} & \cellcolor{gray!6}{30.0} & \cellcolor{gray!6}{85.0} & \cellcolor{gray!6}{101.0} & \cellcolor{gray!6}{121.0}\\
\hline
Mean income & Anaheim & 127.0 & 61.0 & 97.0 & 111.0 & 135.0 & 131.0 & 54.0 & 99.0 & 118.0 & 146.0\\
\hline
\hline
\cellcolor{gray!6}{Mean income} & \cellcolor{gray!6}{Covina \& West Covina} & \cellcolor{gray!6}{126.0} & \cellcolor{gray!6}{44.0} & \cellcolor{gray!6}{95.0} & \cellcolor{gray!6}{116.0} & \cellcolor{gray!6}{144.0} & \cellcolor{gray!6}{125.0} & \cellcolor{gray!6}{42.0} & \cellcolor{gray!6}{100.0} & \cellcolor{gray!6}{115.0} & \cellcolor{gray!6}{143.0}\\
\hline
Mean income & East Los Angeles & 89.0 & 32.0 & 72.0 & 83.0 & 100.0 & 97.0 & 33.0 & 77.0 & 93.0 & 106.0\\
\hline
\cellcolor{gray!6}{Mean income} & \cellcolor{gray!6}{Glendale and Pasadena} & \cellcolor{gray!6}{124.0} & \cellcolor{gray!6}{56.0} & \cellcolor{gray!6}{86.0} & \cellcolor{gray!6}{108.0} & \cellcolor{gray!6}{138.0} & \cellcolor{gray!6}{135.0} & \cellcolor{gray!6}{57.0} & \cellcolor{gray!6}{96.0} & \cellcolor{gray!6}{121.0} & \cellcolor{gray!6}{157.0}\\
\hline
Mean income & Hollywood \& Wilshire & 99.0 & 69.0 & 60.0 & 78.0 & 111.0 & 118.0 & 68.0 & 73.0 & 98.0 & 143.0\\
\hline
\cellcolor{gray!6}{Mean income} & \cellcolor{gray!6}{Long Beach} & \cellcolor{gray!6}{106.0} & \cellcolor{gray!6}{49.0} & \cellcolor{gray!6}{72.0} & \cellcolor{gray!6}{102.0} & \cellcolor{gray!6}{129.0} & \cellcolor{gray!6}{113.0} & \cellcolor{gray!6}{45.0} & \cellcolor{gray!6}{80.0} & \cellcolor{gray!6}{104.0} & \cellcolor{gray!6}{137.0}\\
\hline
Mean income & Riverside, San Bernardino, \& Ontario & 107.0 & 41.0 & 79.0 & 99.0 & 128.0 & 114.0 & 42.0 & 83.0 & 103.0 & 137.0\\
\hline
\cellcolor{gray!6}{Mean income} & \cellcolor{gray!6}{San Fernando Valley (East)} & \cellcolor{gray!6}{133.0} & \cellcolor{gray!6}{105.0} & \cellcolor{gray!6}{78.0} & \cellcolor{gray!6}{97.0} & \cellcolor{gray!6}{128.0} & \cellcolor{gray!6}{137.0} & \cellcolor{gray!6}{89.0} & \cellcolor{gray!6}{84.0} & \cellcolor{gray!6}{109.0} & \cellcolor{gray!6}{154.0}\\
\hline
Mean income & San Fernando Valley (West) & 154.0 & 69.0 & 106.0 & 134.0 & 183.0 & 152.0 & 65.0 & 105.0 & 138.0 & 180.0\\
\hline
\cellcolor{gray!6}{Mean income} & \cellcolor{gray!6}{Santa Ana} & \cellcolor{gray!6}{145.0} & \cellcolor{gray!6}{64.0} & \cellcolor{gray!6}{102.0} & \cellcolor{gray!6}{133.0} & \cellcolor{gray!6}{167.0} & \cellcolor{gray!6}{152.0} & \cellcolor{gray!6}{62.0} & \cellcolor{gray!6}{108.0} & \cellcolor{gray!6}{139.0} & \cellcolor{gray!6}{174.0}\\
\hline
Mean income & South \& South Central Los Angeles & 77.0 & 26.0 & 62.0 & 72.0 & 87.0 & 86.0 & 32.0 & 66.0 & 78.0 & 95.0\\
\hline
\cellcolor{gray!6}{Mean income} & \cellcolor{gray!6}{South Bay} & \cellcolor{gray!6}{141.0} & \cellcolor{gray!6}{78.0} & \cellcolor{gray!6}{89.0} & \cellcolor{gray!6}{118.0} & \cellcolor{gray!6}{164.0} & \cellcolor{gray!6}{149.0} & \cellcolor{gray!6}{77.0} & \cellcolor{gray!6}{95.0} & \cellcolor{gray!6}{124.0} & \cellcolor{gray!6}{180.0}\\
\hline
Mean income & South Orange Co & 173.0 & 66.0 & 125.0 & 163.0 & 203.0 & 179.0 & 62.0 & 138.0 & 168.0 & 213.0\\
\hline
\cellcolor{gray!6}{Mean income} & \cellcolor{gray!6}{West Los Angeles \& West Beach} & \cellcolor{gray!6}{197.0} & \cellcolor{gray!6}{136.0} & \cellcolor{gray!6}{108.0} & \cellcolor{gray!6}{144.0} & \cellcolor{gray!6}{266.0} & \cellcolor{gray!6}{206.0} & \cellcolor{gray!6}{108.0} & \cellcolor{gray!6}{133.0} & \cellcolor{gray!6}{169.0} & \cellcolor{gray!6}{252.0}\\
\hline
Mean income & Whittier \& North Orange Co & 121.0 & 46.0 & 96.0 & 110.0 & 135.0 & 125.0 & 37.0 & 102.0 & 117.0 & 142.0\\
\hline
\cellcolor{gray!6}{\% in poverty} & \cellcolor{gray!6}{Anaheim} & \cellcolor{gray!6}{9.8} & \cellcolor{gray!6}{6.9} & \cellcolor{gray!6}{4.6} & \cellcolor{gray!6}{8.0} & \cellcolor{gray!6}{13.6} & \cellcolor{gray!6}{10.7} & \cellcolor{gray!6}{6.7} & \cellcolor{gray!6}{5.5} & \cellcolor{gray!6}{9.6} & \cellcolor{gray!6}{14.8}\\
\hline
\hline
\% in poverty & Covina \& West Covina & 10.3 & 6.7 & 5.2 & 8.8 & 13.5 & 10.3 & 6.4 & 5.8 & 9.0 & 12.8\\
\hline
\cellcolor{gray!6}{\% in poverty} & \cellcolor{gray!6}{East Los Angeles} & \cellcolor{gray!6}{19.2} & \cellcolor{gray!6}{9.5} & \cellcolor{gray!6}{11.4} & \cellcolor{gray!6}{18.8} & \cellcolor{gray!6}{25.4} & \cellcolor{gray!6}{16.0} & \cellcolor{gray!6}{7.9} & \cellcolor{gray!6}{10.2} & \cellcolor{gray!6}{15.1} & \cellcolor{gray!6}{21.1}\\
\hline
\% in poverty & Glendale and Pasadena & 12.2 & 7.4 & 6.5 & 10.7 & 16.9 & 12.6 & 7.7 & 6.4 & 11.3 & 16.8\\
\hline
\cellcolor{gray!6}{\% in poverty} & \cellcolor{gray!6}{Hollywood \& Wilshire} & \cellcolor{gray!6}{22.1} & \cellcolor{gray!6}{12.9} & \cellcolor{gray!6}{11.5} & \cellcolor{gray!6}{20.8} & \cellcolor{gray!6}{31.0} & \cellcolor{gray!6}{17.7} & \cellcolor{gray!6}{11.8} & \cellcolor{gray!6}{9.3} & \cellcolor{gray!6}{14.7} & \cellcolor{gray!6}{23.9}\\
\hline
\% in poverty & Long Beach & 16.4 & 13.1 & 5.5 & 12.5 & 25.1 & 14.5 & 12.0 & 6.8 & 12.1 & 19.9\\
\hline
\cellcolor{gray!6}{\% in poverty} & \cellcolor{gray!6}{Riverside, San Bernardino, \& Ontario} & \cellcolor{gray!6}{13.0} & \cellcolor{gray!6}{9.8} & \cellcolor{gray!6}{5.7} & \cellcolor{gray!6}{10.8} & \cellcolor{gray!6}{17.0} & \cellcolor{gray!6}{12.4} & \cellcolor{gray!6}{9.2} & \cellcolor{gray!6}{5.8} & \cellcolor{gray!6}{10.6} & \cellcolor{gray!6}{17.2}\\
\hline
\% in poverty & San Fernando Valley (East) & 14.6 & 9.0 & 7.4 & 12.7 & 19.5 & 14.6 & 8.9 & 8.2 & 13.3 & 19.0\\
\hline
\cellcolor{gray!6}{\% in poverty} & \cellcolor{gray!6}{San Fernando Valley (West)} & \cellcolor{gray!6}{8.2} & \cellcolor{gray!6}{7.0} & \cellcolor{gray!6}{3.5} & \cellcolor{gray!6}{6.0} & \cellcolor{gray!6}{10.8} & \cellcolor{gray!6}{9.8} & \cellcolor{gray!6}{7.8} & \cellcolor{gray!6}{4.2} & \cellcolor{gray!6}{8.1} & \cellcolor{gray!6}{13.3}\\
\hline
\% in poverty & Santa Ana & 8.7 & 7.1 & 3.6 & 6.3 & 12.2 & 9.8 & 6.4 & 5.0 & 8.3 & 13.0\\
\hline
\cellcolor{gray!6}{\% in poverty} & \cellcolor{gray!6}{South \& South Central Los Angeles} & \cellcolor{gray!6}{26.7} & \cellcolor{gray!6}{12.7} & \cellcolor{gray!6}{17.9} & \cellcolor{gray!6}{26.1} & \cellcolor{gray!6}{35.7} & \cellcolor{gray!6}{21.7} & \cellcolor{gray!6}{11.5} & \cellcolor{gray!6}{13.7} & \cellcolor{gray!6}{20.6} & \cellcolor{gray!6}{28.5}\\
\hline
\% in poverty & South Bay & 10.2 & 7.9 & 3.9 & 7.2 & 16.4 & 9.6 & 6.9 & 4.9 & 7.5 & 13.3\\
\hline
\cellcolor{gray!6}{\% in poverty} & \cellcolor{gray!6}{South Orange Co} & \cellcolor{gray!6}{5.4} & \cellcolor{gray!6}{5.9} & \cellcolor{gray!6}{2.5} & \cellcolor{gray!6}{4.0} & \cellcolor{gray!6}{6.4} & \cellcolor{gray!6}{7.5} & \cellcolor{gray!6}{6.0} & \cellcolor{gray!6}{3.7} & \cellcolor{gray!6}{6.1} & \cellcolor{gray!6}{9.6}\\
\hline
\% in poverty & West Los Angeles \& West Beach & 11.3 & 12.0 & 5.0 & 8.3 & 13.2 & 10.4 & 10.5 & 5.2 & 8.2 & 11.8\\
\hline
\cellcolor{gray!6}{\% in poverty} & \cellcolor{gray!6}{Whittier \& North Orange Co} & \cellcolor{gray!6}{9.3} & \cellcolor{gray!6}{5.7} & \cellcolor{gray!6}{5.1} & \cellcolor{gray!6}{8.1} & \cellcolor{gray!6}{12.3} & \cellcolor{gray!6}{9.3} & \cellcolor{gray!6}{5.9} & \cellcolor{gray!6}{4.9} & \cellcolor{gray!6}{8.3} & \cellcolor{gray!6}{12.8}\\
\hline
\% with BA+ & Anaheim & 22.2 & 15.5 & 11.9 & 16.5 & 26.3 & 31.1 & 17.4 & 19.2 & 25.3 & 40.8\\
\hline
\hline
\cellcolor{gray!6}{\% with BA+} & \cellcolor{gray!6}{Covina \& West Covina} & \cellcolor{gray!6}{23.8} & \cellcolor{gray!6}{16.0} & \cellcolor{gray!6}{10.2} & \cellcolor{gray!6}{20.4} & \cellcolor{gray!6}{35.3} & \cellcolor{gray!6}{30.5} & \cellcolor{gray!6}{16.6} & \cellcolor{gray!6}{16.2} & \cellcolor{gray!6}{28.5} & \cellcolor{gray!6}{44.0}\\
\hline
\% with BA+ & East Los Angeles & 13.6 & 12.8 & 4.6 & 8.3 & 20.5 & 21.2 & 13.5 & 11.5 & 16.2 & 30.1\\
\hline
\cellcolor{gray!6}{\% with BA+} & \cellcolor{gray!6}{Glendale and Pasadena} & \cellcolor{gray!6}{29.4} & \cellcolor{gray!6}{17.6} & \cellcolor{gray!6}{16.2} & \cellcolor{gray!6}{27.5} & \cellcolor{gray!6}{40.5} & \cellcolor{gray!6}{39.1} & \cellcolor{gray!6}{19.3} & \cellcolor{gray!6}{24.2} & \cellcolor{gray!6}{39.2} & \cellcolor{gray!6}{53.7}\\
\hline
\% with BA+ & Hollywood \& Wilshire & 25.5 & 18.5 & 10.0 & 19.6 & 42.2 & 42.1 & 19.9 & 25.2 & 40.6 & 59.8\\
\hline
\cellcolor{gray!6}{\% with BA+} & \cellcolor{gray!6}{Long Beach} & \cellcolor{gray!6}{20.9} & \cellcolor{gray!6}{16.2} & \cellcolor{gray!6}{7.1} & \cellcolor{gray!6}{16.9} & \cellcolor{gray!6}{31.7} & \cellcolor{gray!6}{28.4} & \cellcolor{gray!6}{17.1} & \cellcolor{gray!6}{13.6} & \cellcolor{gray!6}{25.9} & \cellcolor{gray!6}{40.8}\\
\hline
\% with BA+ & Riverside, San Bernardino, \& Ontario & 16.6 & 12.7 & 7.0 & 12.4 & 23.0 & 22.5 & 14.2 & 10.8 & 19.4 & 31.3\\
\hline
\cellcolor{gray!6}{\% with BA+} & \cellcolor{gray!6}{San Fernando Valley (East)} & \cellcolor{gray!6}{28.2} & \cellcolor{gray!6}{18.6} & \cellcolor{gray!6}{13.7} & \cellcolor{gray!6}{23.3} & \cellcolor{gray!6}{40.3} & \cellcolor{gray!6}{39.3} & \cellcolor{gray!6}{21.4} & \cellcolor{gray!6}{21.3} & \cellcolor{gray!6}{34.4} & \cellcolor{gray!6}{59.1}\\
\hline
\% with BA+ & San Fernando Valley (West) & 30.1 & 16.3 & 18.2 & 27.7 & 43.2 & 38.0 & 17.8 & 24.7 & 37.1 & 52.0\\
\hline
\cellcolor{gray!6}{\% with BA+} & \cellcolor{gray!6}{Santa Ana} & \cellcolor{gray!6}{30.6} & \cellcolor{gray!6}{19.4} & \cellcolor{gray!6}{12.9} & \cellcolor{gray!6}{31.2} & \cellcolor{gray!6}{46.1} & \cellcolor{gray!6}{40.6} & \cellcolor{gray!6}{20.4} & \cellcolor{gray!6}{23.8} & \cellcolor{gray!6}{40.8} & \cellcolor{gray!6}{56.3}\\
\hline
\% with BA+ & South \& South Central Los Angeles & 8.9 & 10.0 & 3.1 & 5.4 & 10.1 & 16.2 & 14.9 & 7.1 & 10.8 & 19.5\\
\hline
\cellcolor{gray!6}{\% with BA+} & \cellcolor{gray!6}{South Bay} & \cellcolor{gray!6}{31.7} & \cellcolor{gray!6}{20.8} & \cellcolor{gray!6}{13.7} & \cellcolor{gray!6}{25.8} & \cellcolor{gray!6}{46.9} & \cellcolor{gray!6}{41.5} & \cellcolor{gray!6}{21.8} & \cellcolor{gray!6}{22.6} & \cellcolor{gray!6}{36.7} & \cellcolor{gray!6}{58.8}\\
\hline
\% with BA+ & South Orange Co & 40.4 & 16.3 & 29.2 & 40.0 & 53.4 & 52.3 & 15.7 & 42.9 & 51.6 & 63.8\\
\hline
\cellcolor{gray!6}{\% with BA+} & \cellcolor{gray!6}{West Los Angeles \& West Beach} & \cellcolor{gray!6}{50.6} & \cellcolor{gray!6}{16.7} & \cellcolor{gray!6}{42.3} & \cellcolor{gray!6}{53.4} & \cellcolor{gray!6}{62.7} & \cellcolor{gray!6}{65.6} & \cellcolor{gray!6}{14.4} & \cellcolor{gray!6}{60.1} & \cellcolor{gray!6}{67.9} & \cellcolor{gray!6}{74.6}\\
\hline
\% with BA+ & Whittier \& North Orange Co & 21.4 & 14.0 & 10.2 & 16.3 & 32.0 & 29.0 & 14.9 & 17.9 & 23.9 & 40.9\\
\hline
\end{tabular}
\end{table}