### 2024-10-18 aefp/results_test.qmd

---
title: "Analysis Results"
output: 
  pdf_document:
    keep_tex: true
header-includes:
  - \usepackage{booktabs}
  - \usepackage{colortbl}
  - \usepackage[table]{xcolor}
  - \usepackage{array} # For custom column types if needed
---

## Results of the Analysis

Here we present the summarized data for EPS variables across different years.

```{r table-inclusion, echo=FALSE, results='asis'}
#cat(readLines("eps_table.tex"), sep = "\n")
getwd()
#cat(readLines(file.path('.','..','results','tables',"socal_table.tex")), sep = "\n")
#\input{../results/tables/socal_table.tex}
```
\input{../results/tables/socal_table.tex}