### 2024-10-18 results/tables/socal_table.tex

\begingroup\fontsize{10}{12}\selectfont

\begin{longtable}[l]{l|rrrrr}
\caption{Characteristics of Geomarkets around Los Angeles {#tab:socal_eps_table_summary}}\\
\toprule
\multicolumn{1}{c}{ } & \multicolumn{5}{c}{2020} \\
\textbf{Geomarket name} & \textbf{Mean} & \textbf{SD} & \textbf{P25} & \textbf{P50} & \textbf{P75}\\
\midrule
\endfirsthead
\caption[]{Characteristics of Geomarkets around Los Angeles {#tab:socal_eps_table_summary} \textit{(continued)}}\\
\toprule
\textbf{Geomarket name} & \textbf{Mean} & \textbf{SD} & \textbf{P25} & \textbf{P50} & \textbf{P75}\\
\midrule
\endhead

\endfoot
\bottomrule
\endlastfoot
\addlinespace[0.5em]
\multicolumn{6}{l}{\textbf{\% White, non-Hispanic}}\\
\hspace{1em}CA14, San Fernando Valley (West) & 44.0 & 24.2 & 22.8 & 46.6 & 64.7\\
\hspace{1em}CA15, San Fernando Valley (East) & 43.3 & 25.7 & 22.5 & 39.6 & 67.0\\
\hspace{1em}CA16, Glendale and Pasadena & 35.3 & 24.0 & 12.8 & 34.1 & 57.2\\
\hspace{1em}CA17, West Los Angeles & West Beach & 60.2 & 20.2 & 45.9 & 62.9 & 75.5\\
\hspace{1em}CA18, Hollywood & Wilshire & 31.2 & 25.7 & 8.1 & 22.6 & 53.5\\
\hspace{1em}CA19, East Los Angeles & 6.5 & 7.1 & 2.0 & 4.1 & 8.4\\
\hspace{1em}CA20, South Bay & 29.1 & 24.5 & 7.1 & 22.1 & 51.4\\
\hspace{1em}CA21, South & South Central Los Angeles & 6.3 & 11.4 & 0.6 & 1.9 & 6.6\\
\hspace{1em}CA22, Long Beach & 25.8 & 23.0 & 6.6 & 16.4 & 45.5\\
\hspace{1em}CA23, Covina & West Covina & 18.8 & 16.8 & 5.4 & 12.5 & 27.1\\
\hspace{1em}CA24, Whittier & North Orange Co & 22.4 & 15.3 & 10.9 & 17.5 & 32.7\\
\hspace{1em}CA25, Anaheim & 26.1 & 17.8 & 12.6 & 20.4 & 36.9\\
\hspace{1em}CA26, Santa Ana & 42.0 & 26.3 & 17.0 & 44.1 & 64.6\\
\hspace{1em}CA27, Riverside, San Bernardino, & Ontario & 27.7 & 18.8 & 12.5 & 23.5 & 40.7\\
\hspace{1em}CA28, South Orange Co & 52.9 & 18.1 & 40.7 & 55.0 & 66.3\\
\addlinespace[0.5em]
\multicolumn{6}{l}{\textbf{\% Black, non-Hispanic}}\\
\hspace{1em}CA14, San Fernando Valley (West) & 3.8 & 3.9 & 1.1 & 2.7 & 5.4\\
\hspace{1em}CA15, San Fernando Valley (East) & 4.3 & 3.8 & 1.6 & 3.3 & 5.7\\
\hspace{1em}CA16, Glendale and Pasadena & 3.4 & 5.2 & 0.5 & 1.8 & 4.1\\
\hspace{1em}CA17, West Los Angeles & West Beach & 5.4 & 8.1 & 1.1 & 2.9 & 6.2\\
\hspace{1em}CA18, Hollywood & Wilshire & 11.1 & 16.8 & 1.9 & 4.5 & 10.5\\
\hspace{1em}CA19, East Los Angeles & 1.3 & 2.8 & 0.0 & 0.4 & 1.4\\
\hspace{1em}CA20, South Bay & 12.1 & 16.9 & 1.7 & 5.7 & 14.3\\
\hspace{1em}CA21, South & South Central Los Angeles & 17.5 & 18.8 & 2.7 & 11.6 & 25.3\\
\hspace{1em}CA22, Long Beach & 11.4 & 13.0 & 3.6 & 8.1 & 15.1\\
\hspace{1em}CA23, Covina & West Covina & 3.2 & 3.1 & 0.8 & 2.3 & 4.7\\
\hspace{1em}CA24, Whittier & North Orange Co & 2.4 & 2.7 & 0.4 & 1.5 & 3.0\\
\hspace{1em}CA25, Anaheim & 4.4 & 5.6 & 0.8 & 2.3 & 6.0\\
\hspace{1em}CA26, Santa Ana & 1.3 & 1.9 & 0.1 & 0.7 & 1.6\\
\hspace{1em}CA27, Riverside, San Bernardino, & Ontario & 7.1 & 6.4 & 2.4 & 5.4 & 9.8\\
\hspace{1em}CA28, South Orange Co & 1.5 & 1.8 & 0.3 & 1.0 & 2.3\\
\addlinespace[0.5em]
\multicolumn{6}{l}{\textbf{\% Hispanic}}\\
\hspace{1em}CA14, San Fernando Valley (West) & 36.8 & 25.9 & 16.2 & 28.8 & 56.9\\
\hspace{1em}CA15, San Fernando Valley (East) & 40.6 & 27.2 & 14.8 & 37.3 & 60.9\\
\hspace{1em}CA16, Glendale and Pasadena & 36.8 & 26.9 & 16.3 & 26.3 & 53.4\\
\hspace{1em}CA17, West Los Angeles & West Beach & 15.8 & 13.4 & 6.3 & 12.2 & 20.4\\
\hspace{1em}CA18, Hollywood & Wilshire & 38.2 & 24.6 & 15.7 & 34.7 & 57.1\\
\hspace{1em}CA19, East Los Angeles & 64.4 & 28.9 & 36.7 & 73.2 & 91.5\\
\hspace{1em}CA20, South Bay & 33.2 & 22.5 & 12.9 & 29.8 & 51.8\\
\hspace{1em}CA21, South & South Central Los Angeles & 70.5 & 24.8 & 60.2 & 75.7 & 90.2\\
\hspace{1em}CA22, Long Beach & 45.8 & 23.3 & 26.6 & 43.5 & 63.4\\
\hspace{1em}CA23, Covina & West Covina & 51.7 & 23.5 & 30.8 & 54.3 & 71.3\\
\hspace{1em}CA24, Whittier & North Orange Co & 56.0 & 23.6 & 39.3 & 59.2 & 75.7\\
\hspace{1em}CA25, Anaheim & 41.3 & 22.7 & 23.1 & 38.6 & 58.1\\
\hspace{1em}CA26, Santa Ana & 33.0 & 27.9 & 12.3 & 21.2 & 46.6\\
\hspace{1em}CA27, Riverside, San Bernardino, & Ontario & 53.8 & 22.0 & 36.3 & 54.1 & 72.4\\
\hspace{1em}CA28, South Orange Co & 22.2 & 17.0 & 10.4 & 16.8 & 27.2\\
\addlinespace[0.5em]
\multicolumn{6}{l}{\textbf{\% API, non-Hispanic}}\\
\hspace{1em}CA14, San Fernando Valley (West) & 11.6 & 8.1 & 5.3 & 10.4 & 16.3\\
\hspace{1em}CA15, San Fernando Valley (East) & 8.3 & 5.4 & 4.7 & 6.9 & 10.2\\
\hspace{1em}CA16, Glendale and Pasadena & 20.9 & 19.8 & 7.1 & 13.0 & 27.3\\
\hspace{1em}CA17, West Los Angeles & West Beach & 13.1 & 9.5 & 6.5 & 9.8 & 17.6\\
\hspace{1em}CA18, Hollywood & Wilshire & 15.7 & 14.0 & 5.7 & 11.4 & 20.1\\
\hspace{1em}CA19, East Los Angeles & 26.5 & 26.0 & 2.7 & 16.4 & 47.9\\
\hspace{1em}CA20, South Bay & 21.2 & 15.9 & 8.2 & 17.8 & 30.3\\
\hspace{1em}CA21, South & South Central Los Angeles & 3.9 & 6.9 & 0.2 & 1.2 & 4.4\\
\hspace{1em}CA22, Long Beach & 13.2 & 11.3 & 4.9 & 10.2 & 18.2\\
\hspace{1em}CA23, Covina & West Covina & 23.7 & 19.3 & 10.0 & 16.7 & 31.9\\
\hspace{1em}CA24, Whittier & North Orange Co & 15.9 & 16.2 & 5.2 & 12.0 & 18.0\\
\hspace{1em}CA25, Anaheim & 24.5 & 18.1 & 11.2 & 19.4 & 34.2\\
\hspace{1em}CA26, Santa Ana & 20.1 & 18.3 & 6.5 & 14.3 & 26.5\\
\hspace{1em}CA27, Riverside, San Bernardino, & Ontario & 8.2 & 10.0 & 2.0 & 5.0 & 10.4\\
\hspace{1em}CA28, South Orange Co & 18.8 & 15.1 & 9.0 & 14.1 & 23.6\\
\addlinespace[0.5em]
\multicolumn{6}{l}{\textbf{\% Native, non-Hispanic}}\\
\hspace{1em}CA14, San Fernando Valley (West) & 0.2 & 0.5 & 0.0 & 0.0 & 0.1\\
\hspace{1em}CA15, San Fernando Valley (East) & 0.1 & 0.3 & 0.0 & 0.0 & 0.0\\
\hspace{1em}CA16, Glendale and Pasadena & 0.2 & 0.5 & 0.0 & 0.0 & 0.1\\
\hspace{1em}CA17, West Los Angeles & West Beach & 0.2 & 0.4 & 0.0 & 0.0 & 0.0\\
\hspace{1em}CA18, Hollywood & Wilshire & 0.2 & 0.5 & 0.0 & 0.0 & 0.0\\
\hspace{1em}CA19, East Los Angeles & 0.2 & 0.6 & 0.0 & 0.0 & 0.2\\
\hspace{1em}CA20, South Bay & 0.2 & 0.5 & 0.0 & 0.0 & 0.0\\
\hspace{1em}CA21, South & South Central Los Angeles & 0.1 & 0.4 & 0.0 & 0.0 & 0.0\\
\hspace{1em}CA22, Long Beach & 0.6 & 6.9 & 0.0 & 0.0 & 0.1\\
\hspace{1em}CA23, Covina & West Covina & 0.2 & 0.7 & 0.0 & 0.0 & 0.2\\
\hspace{1em}CA24, Whittier & North Orange Co & 0.8 & 7.7 & 0.0 & 0.0 & 0.3\\
\hspace{1em}CA25, Anaheim & 0.5 & 6.0 & 0.0 & 0.0 & 0.2\\
\hspace{1em}CA26, Santa Ana & 0.2 & 0.4 & 0.0 & 0.0 & 0.2\\
\hspace{1em}CA27, Riverside, San Bernardino, & Ontario & 0.3 & 0.6 & 0.0 & 0.0 & 0.3\\
\hspace{1em}CA28, South Orange Co & 0.2 & 0.4 & 0.0 & 0.0 & 0.0\\
\addlinespace[0.5em]
\multicolumn{6}{l}{\textbf{\% two+ races, non-Hispanic}}\\
\hspace{1em}CA14, San Fernando Valley (West) & 3.1 & 2.6 & 1.1 & 2.7 & 4.1\\
\hspace{1em}CA15, San Fernando Valley (East) & 3.1 & 2.9 & 1.0 & 2.7 & 4.4\\
\hspace{1em}CA16, Glendale and Pasadena & 3.0 & 2.7 & 1.0 & 2.4 & 4.2\\
\hspace{1em}CA17, West Los Angeles & West Beach & 4.8 & 2.9 & 2.7 & 4.2 & 6.7\\
\hspace{1em}CA18, Hollywood & Wilshire & 3.2 & 2.9 & 1.3 & 2.4 & 4.6\\
\hspace{1em}CA19, East Los Angeles & 0.9 & 1.2 & 0.0 & 0.4 & 1.4\\
\hspace{1em}CA20, South Bay & 3.8 & 2.9 & 1.5 & 3.2 & 5.6\\
\hspace{1em}CA21, South & South Central Los Angeles & 1.3 & 2.0 & 0.0 & 0.5 & 1.8\\
\hspace{1em}CA22, Long Beach & 2.9 & 2.3 & 1.2 & 2.5 & 4.2\\
\hspace{1em}CA23, Covina & West Covina & 2.0 & 2.1 & 0.4 & 1.5 & 2.9\\
\hspace{1em}CA24, Whittier & North Orange Co & 1.9 & 1.7 & 0.5 & 1.5 & 2.8\\
\hspace{1em}CA25, Anaheim & 2.9 & 2.3 & 1.1 & 2.1 & 4.5\\
\hspace{1em}CA26, Santa Ana & 3.2 & 2.9 & 1.1 & 2.6 & 4.9\\
\hspace{1em}CA27, Riverside, San Bernardino, & Ontario & 2.6 & 2.4 & 0.7 & 1.9 & 3.7\\
\hspace{1em}CA28, South Orange Co & 4.1 & 2.7 & 2.1 & 3.7 & 5.5\\
\addlinespace[0.5em]
\multicolumn{6}{l}{\textbf{Median income}}\\
\hspace{1em}CA14, San Fernando Valley (West) & 121.0 & 46.0 & 87.0 & 116.0 & 146.0\\
\hspace{1em}CA15, San Fernando Valley (East) & 99.0 & 56.0 & 61.0 & 83.0 & 116.0\\
\hspace{1em}CA16, Glendale and Pasadena & 102.0 & 40.0 & 73.0 & 96.0 & 121.0\\
\hspace{1em}CA17, West Los Angeles & West Beach & 137.0 & 61.0 & 98.0 & 122.0 & 165.0\\
\hspace{1em}CA18, Hollywood & Wilshire & 84.0 & 43.0 & 52.0 & 75.0 & 105.0\\
\hspace{1em}CA19, East Los Angeles & 76.0 & 25.0 & 60.0 & 72.0 & 87.0\\
\hspace{1em}CA20, South Bay & 115.0 & 51.0 & 78.0 & 103.0 & 141.0\\
\hspace{1em}CA21, South & South Central Los Angeles & 67.0 & 27.0 & 50.0 & 62.0 & 78.0\\
\hspace{1em}CA22, Long Beach & 90.0 & 36.0 & 62.0 & 85.0 & 111.0\\
\hspace{1em}CA23, Covina & West Covina & 103.0 & 32.0 & 82.0 & 99.0 & 120.0\\
\hspace{1em}CA24, Whittier & North Orange Co & 105.0 & 30.0 & 85.0 & 101.0 & 121.0\\
\hspace{1em}CA25, Anaheim & 108.0 & 41.0 & 78.0 & 99.0 & 129.0\\
\hspace{1em}CA26, Santa Ana & 119.0 & 44.0 & 87.0 & 112.0 & 142.0\\
\hspace{1em}CA27, Riverside, San Bernardino, & Ontario & 95.0 & 37.0 & 68.0 & 88.0 & 119.0\\
\hspace{1em}CA28, South Orange Co & 142.0 & 46.0 & 109.0 & 140.0 & 171.0\\
\addlinespace[0.5em]
\multicolumn{6}{l}{\textbf{Mean income}}\\
\hspace{1em}CA14, San Fernando Valley (West) & 152.0 & 65.0 & 105.0 & 138.0 & 180.0\\
\hspace{1em}CA15, San Fernando Valley (East) & 137.0 & 89.0 & 84.0 & 109.0 & 154.0\\
\hspace{1em}CA16, Glendale and Pasadena & 135.0 & 57.0 & 96.0 & 121.0 & 157.0\\
\hspace{1em}CA17, West Los Angeles & West Beach & 206.0 & 108.0 & 133.0 & 169.0 & 252.0\\
\hspace{1em}CA18, Hollywood & Wilshire & 118.0 & 68.0 & 73.0 & 98.0 & 143.0\\
\hspace{1em}CA19, East Los Angeles & 97.0 & 33.0 & 77.0 & 93.0 & 106.0\\
\hspace{1em}CA20, South Bay & 149.0 & 77.0 & 95.0 & 124.0 & 180.0\\
\hspace{1em}CA21, South & South Central Los Angeles & 86.0 & 32.0 & 66.0 & 78.0 & 95.0\\
\hspace{1em}CA22, Long Beach & 113.0 & 45.0 & 80.0 & 104.0 & 137.0\\
\hspace{1em}CA23, Covina & West Covina & 125.0 & 42.0 & 100.0 & 115.0 & 143.0\\
\hspace{1em}CA24, Whittier & North Orange Co & 125.0 & 37.0 & 102.0 & 117.0 & 142.0\\
\hspace{1em}CA25, Anaheim & 131.0 & 54.0 & 99.0 & 118.0 & 146.0\\
\hspace{1em}CA26, Santa Ana & 152.0 & 62.0 & 108.0 & 139.0 & 174.0\\
\hspace{1em}CA27, Riverside, San Bernardino, & Ontario & 114.0 & 42.0 & 83.0 & 103.0 & 137.0\\
\hspace{1em}CA28, South Orange Co & 179.0 & 62.0 & 138.0 & 168.0 & 213.0\\
\addlinespace[0.5em]
\multicolumn{6}{l}{\textbf{\% in poverty}}\\
\hspace{1em}CA14, San Fernando Valley (West) & 9.8 & 7.8 & 4.2 & 8.1 & 13.3\\
\hspace{1em}CA15, San Fernando Valley (East) & 14.6 & 8.9 & 8.2 & 13.3 & 19.0\\
\hspace{1em}CA16, Glendale and Pasadena & 12.6 & 7.7 & 6.4 & 11.3 & 16.8\\
\hspace{1em}CA17, West Los Angeles & West Beach & 10.4 & 10.5 & 5.2 & 8.2 & 11.8\\
\hspace{1em}CA18, Hollywood & Wilshire & 17.7 & 11.8 & 9.3 & 14.7 & 23.9\\
\hspace{1em}CA19, East Los Angeles & 16.0 & 7.9 & 10.2 & 15.1 & 21.1\\
\hspace{1em}CA20, South Bay & 9.6 & 6.9 & 4.9 & 7.5 & 13.3\\
\hspace{1em}CA21, South & South Central Los Angeles & 21.7 & 11.5 & 13.7 & 20.6 & 28.5\\
\hspace{1em}CA22, Long Beach & 14.5 & 12.0 & 6.8 & 12.1 & 19.9\\
\hspace{1em}CA23, Covina & West Covina & 10.3 & 6.4 & 5.8 & 9.0 & 12.8\\
\hspace{1em}CA24, Whittier & North Orange Co & 9.3 & 5.9 & 4.9 & 8.3 & 12.8\\
\hspace{1em}CA25, Anaheim & 10.7 & 6.7 & 5.5 & 9.6 & 14.8\\
\hspace{1em}CA26, Santa Ana & 9.8 & 6.4 & 5.0 & 8.3 & 13.0\\
\hspace{1em}CA27, Riverside, San Bernardino, & Ontario & 12.4 & 9.2 & 5.8 & 10.6 & 17.2\\
\hspace{1em}CA28, South Orange Co & 7.5 & 6.0 & 3.7 & 6.1 & 9.6\\
\addlinespace[0.5em]
\multicolumn{6}{l}{\textbf{\% with BA+}}\\
\hspace{1em}CA14, San Fernando Valley (West) & 38.0 & 17.8 & 24.7 & 37.1 & 52.0\\
\hspace{1em}CA15, San Fernando Valley (East) & 39.3 & 21.4 & 21.3 & 34.4 & 59.1\\
\hspace{1em}CA16, Glendale and Pasadena & 39.1 & 19.3 & 24.2 & 39.2 & 53.7\\
\hspace{1em}CA17, West Los Angeles & West Beach & 65.6 & 14.4 & 60.1 & 67.9 & 74.6\\
\hspace{1em}CA18, Hollywood & Wilshire & 42.1 & 19.9 & 25.2 & 40.6 & 59.8\\
\hspace{1em}CA19, East Los Angeles & 21.2 & 13.5 & 11.5 & 16.2 & 30.1\\
\hspace{1em}CA20, South Bay & 41.5 & 21.8 & 22.6 & 36.7 & 58.8\\
\hspace{1em}CA21, South & South Central Los Angeles & 16.2 & 14.9 & 7.1 & 10.8 & 19.5\\
\hspace{1em}CA22, Long Beach & 28.4 & 17.1 & 13.6 & 25.9 & 40.8\\
\hspace{1em}CA23, Covina & West Covina & 30.5 & 16.6 & 16.2 & 28.5 & 44.0\\
\hspace{1em}CA24, Whittier & North Orange Co & 29.0 & 14.9 & 17.9 & 23.9 & 40.9\\
\hspace{1em}CA25, Anaheim & 31.1 & 17.4 & 19.2 & 25.3 & 40.8\\
\hspace{1em}CA26, Santa Ana & 40.6 & 20.4 & 23.8 & 40.8 & 56.3\\
\hspace{1em}CA27, Riverside, San Bernardino, & Ontario & 22.5 & 14.2 & 10.8 & 19.4 & 31.3\\
\hspace{1em}CA28, South Orange Co & 52.3 & 15.7 & 42.9 & 51.6 & 63.8\\*
\end{longtable}
\endgroup{}