### 2025-01-30 text/cb_geomarket_online_appendix.qmd
tweaker

---
title: "Online appendix"
format:
  pdf:
    toc: true
    toc-depth: 3
    number-sections: true
---

# header 1

```{r}
getwd()
```

```{r}
# You can load libraries, set options, define your list of metros, etc.
library(tools)  # for toTitleCase() if desired

# Example lists
metros <- c("chicago", "denver")
RQs <- c("RQ1", "RQ2")

# Define the function to insert a figure, given a metro + RQ.
# 'base_path' is where your .png and .txt files live.
insert_figure <- function(metro, rq, base_path = "../results/graphs") {

  # Construct file paths
  fig_file    <- file.path(base_path, paste0(metro, "_", rq, ".png"))
  cap_file    <- file.path(base_path, paste0(metro, "_", rq, "_caption.txt"))
  note_file   <- file.path(base_path, paste0(metro, "_", rq, "_note.txt"))

  # Read the caption (if missing, use a default)
  if (file.exists(cap_file)) {
    caption_text <- paste(readLines(cap_file, warn = FALSE), collapse = " ")
  } else {
    caption_text <- paste("No caption found for", metro, rq)
  }

  # Read the note (if missing, it remains empty)
  if (file.exists(note_file)) {
    note_text <- paste(readLines(note_file, warn = FALSE), collapse = " ")
    # Format the note to appear after the figure
    note_string <- paste0("\n\n**Note**: ", note_text)
  } else {
    note_string <- ""
  }

  # Build the Quarto figure block as a string
  # - We use fig-cap= for the caption
  # - We can specify width, alignment, etc.
  # - Insert the note at the bottom if present
  fig_block <- paste0(
    "::: {float='figure' fig-align='center' fig-cap='", caption_text, "'}\n\n",
    "![](", fig_file, "){width='70%'}\n",
    note_string, "\n",
    ":::\n"
  )

  return(fig_block)
}

```