### 2025-01-31 text/cb_geomarket_online_appendix.qmd
creating appendix doc with all graphs

---
title: "Online appendix"
format:
  pdf:
    toc: true
    toc-depth: 3
    number-sections: true
    documentclass: article
    classoption: landscape
    geometry: margin=0.1in
    include-in-header:
      text: |
        \pagenumbering{gobble}    
header-includes:
  - |
    \usepackage{ragged2e}
    \usepackage{caption}
    \captionsetup[figure]{justification=RaggedRight, singlelinecheck=false}    
---

```{r, include = FALSE, eval=FALSE}
getwd()
file.exists("../results/graphs/rq1_chicago_ses.png")

```
<!-- this code run in markdown would include the graph: ![](../results/graphs/rq1_chicago_ses.png){width=70%} -->

```{r, include = FALSE}
# You can load libraries, set options, define your list of metros, etc.
library(tools)  # for toTitleCase() if desired
library(tidyverse)

# Example lists
metros <- c(
  "atlanta",
  "bay area",
  "boston",
  "chicago",
  "cleveland",
  "dallas",
  "dc maryland virginia",
  "detroit",
  "houston",
  "long island",
  "los angeles",
  "miami",
  "new york city",
  "northern new jersey",
  "orange county",
  "philadelphia",
  "san diego"
)

metros <- c(
  "atlanta",
  "chicago",
  "philadelphia"
)

rqs <- c("rq1",'rq2')
graph_types = c('race','ses')

# Define the function to insert a figure, given a metro + RQ.
# 'base_path' is where your .png and .txt files live.
insert_figure <- function(rq, metro, graph_type,base_path = "../results/graphs") {

  metro <- str_replace_all(string = metro, pattern = ' ', replacement = '_')
  # Construct file paths
  fig_file    <- file.path(base_path, str_c(rq,'_',metro,'_',graph_type, ".png"))
  cap_file    <- file.path(base_path, str_c(rq,'_',metro,'_',graph_type,"_title.txt"))
  note_file    <- file.path(base_path, str_c(rq,'_',metro,'_',graph_type,"_note.txt"))

  # Read the caption (if missing, use a default)
  if (file.exists(cap_file)) {
    caption_text <- paste(readLines(cap_file, warn = FALSE), collapse = " ")
  } else {
    caption_text <- paste("No caption found for", metro, rq)
  }

  # Read the note (if missing, it remains empty)
  if (file.exists(note_file)) {
    note_text <- paste(readLines(note_file, warn = FALSE), collapse = " ")
    # Format the note to appear after the figure
    note_string <- paste0("\n**Note**: ", note_text)
  } else {
    note_string <- ""
  }

  # Build the Quarto figure block as a string
  # - We use fig-cap= for the caption
  # - We can specify width, alignment, etc.
  # - Insert the note at the bottom if present
fig_block <- str_c(
  "::: {.figure}\n\n",
  "![", caption_text, "](", fig_file, "){fig-align='left' width='101%'}\n",
  note_string, "\n",
  ":::\n"
)

  return(fig_block)
}

#rq2_philadelphia_order_448922_race_by_firstgen
#rq1_philadelphia_ses
#rq2_philadelphia_firstgen_row_plot

#insert_figure(rq = 'rq2', metro = 'philadelphia', graph_type = 'firstgen_row_plot',base_path = "../results/graphs")
#insert_figure(rq = 'rq2', metro = 'philadelphia', graph_type = 'order_448922_race_by_firstgen',base_path = "../results/graphs")
#insert_figure(rq = 'rq1', metro = 'dallas', graph_type = 'ses',base_path = "../results/graphs")
```

```{r, results = 'asis', echo = FALSE}
# results='asis' so that the text we "cat()" is interpreted as Markdown.

for (m in metros) {

  # Print a top-level heading for the metro area
  cat("\n# ", toTitleCase(m), " area\n", sep="")

  for (rq in rqs) {

    rq_num <- str_replace(string = rq, pattern = "rq", replacement = "")

    if (rq == 'rq1') { # rq1 run for every metro area

      # Print a second-level heading for the RQ
      cat("## Research question ", rq_num, "\n", sep="")

      for (g in graph_types) {

        # Call our figure-inserting function
        figure_md <- insert_figure(metro = m, rq = rq, graph_type = g)

        # Output the figure block to the doc
        cat(figure_md, "\n")
      } # for (g in graph_types) {

    } else if (rq == 'rq2' & (m %in% c('chicago','philadelphia'))) { # rq2 run only for specific metro areas and orders

      # Print a second-level heading for the RQ
      cat("## Research question ", rq_num, "\n", sep="")

      if (m == 'chicago') {

        # Call our figure-inserting function
        figure_md <- insert_figure(metro = m, rq = rq, graph_type = 'order_487984_race_by_firstgen')

        # Output the figure block to the doc
        cat(figure_md, "\n")

        figure_md <- insert_figure(metro = m, rq = rq, graph_type = 'order_488035_488053_race_by_firstgen')

        # Output the figure block to the doc
        cat(figure_md, "\n")

      } else if (m == 'philadelphia') {

        figure_md <- insert_figure(metro = m, rq = rq, graph_type = 'race_row_plot')
        cat(figure_md, "\n")

        figure_md <- insert_figure(metro = m, rq = rq, graph_type = 'race_col_plot')
        cat(figure_md, "\n")

        figure_md <- insert_figure(metro = m, rq = rq, graph_type = 'firstgen_row_plot')
        cat(figure_md, "\n")

        figure_md <- insert_figure(metro = m, rq = rq, graph_type = 'firstgen_col_plot')
        cat(figure_md, "\n")

        # Call our figure-inserting function
        figure_md <- insert_figure(metro = m, rq = rq, graph_type = 'order_448922_race_by_firstgen')

        # Output the figure block to the doc
        cat(figure_md, "\n")

        figure_md <- insert_figure(metro = m, rq = rq, graph_type = 'order_448427_448440_race_by_firstgen')

        # Output the figure block to the doc
        cat(figure_md, "\n")

      } #else if (condition3) {
        # Code block for condition3
      #}

    }

  } # for (rq in rqs) {

}

```