### 2025-04-10 index.qmd
adding capabilities to slides

---
title: 'Structuring College Access'
subtitle: 'The Market Segment Model and College Board Geomarkets'
author: ['Ozan Jaquette', 'Karina Salazar', 'Crystal Han']
institute: ['UCLA', 'University of Arizona', 'Stanford University']
bibliography: ./assets/bib/cb_geomarket_bib.bib
format:
  revealjs:
    revealjs-plugins: [multimodal]   # ← correct key
    theme: [default, ./assets/css/custom.scss]
    controls: true
    controls-layout: bottom-right
    controls-tutorial: true
    transition: slide
    background-transition: fade
    auto-stretch: false
    slide-level: 3
    menu:
      titleSelector: 'h1, h2, .slide:not(.hide) h3'
      useTextContentForMissingTitles: false
      hideMissingTitles: true
    template-partials:
      - ./assets/html/title-slide.html
include-in-header: ./assets/html/header.html
---

# Introduction

### [__College Board _Student Search Service_, order \#448922__](./assets/images/geomarket_orders_redacted.pdf){target="_blank"} {.hide}

__"Geomarkets" carve states and metro areas into smaller geographic regions__

- Geomarkets are an input in two College Board recruiting products:
  - Student Search Service (est. 1972); Enrollment Planning Service (est. 1984)

__Perspectives from sociology on sorting students into colleges:__

- Status attainment model [e.g., @RN915]
- Cultural capital model [e.g., @RN1228]
- Enrollment management behavior of colleges [e.g., @RN3519]
- Sociology ignores third-party products that sort students on behalf of colleges
  - Designed to amplify historic inequality in student demand

<!-- 
__Sociology misses an important mechanism of social reproduction:__

- Third-party vendors/products sort students on behalf of colleges
  - Amplify past structural inequality in student demand
 -->

__Research questions:__

1. What is SES, racial variation between, within Geomarkets? Over time?
1. What is SES and racial composition of included vs. excluded prospects when student list purchases filter on particular Geomarkets?

<!--

### Introduction og {.hide}

Two recruiting products sold by College Board

- Student Search Service (est. 1972): purchase contact information of prospects that meet "search filter" criteria [@RN5012]
- Enrollment Planning Service (EPS, est. 1984): software recommends which "Geomarkets" to target and which high schools to visit

College Board "Geomarkets" are an input in both products

- Geomarkets carve states and metro areas into smaller geographic regions

Concerns: Geomarket borders correlated with race/class; likely to be used in ways that undermine equality of opportunity

1. What is SES, racial variation between, within Geomarkets? Over time?
1. What is SES and racial composition of included vs. excluded prospects when student list purchases filter on particular Geomarkets?

 -->

### Geomarkets around Chicago-land {.hide}

::: columns
::: {.column width="50%"}

![](./assets/images/chicago_geomarkets.png)

<!-- <img src="./assets/images/nyc_geomarkets.png" data-width="60%"> -->
:::

::: {.column width="50%"}
**[University of Chicago admissions counselors, IL territories](https://collegeadmissions.uchicago.edu/contact)**
![](./assets/images/territories/chicagoland_territories_u_chicago_redacted.png)
:::
:::

### Presentation Outline {.hide}

<br>

1. <span style="color:blue">Enrollment management: Background & scholarship</span>  
1. <span style="color:red">Empirical case and theory</span>  
1. <span style="color:red">Methods</span>  
1. <span style="color:red">Results</span>  
1. <span style="color:blue">Discussion<br>– Creating a "knowledge infrastructure" [@RN5025] to study the role of private firms and private equity funding in education</span>

# Enrollment Management: Background \& Scholarship

### __Enrollment Management (EM)__ {.hide .columns-2}

::: columns

:::: {.column width=65%}

__Enrollment management__

- EM profession integrates techniques from marketing and economics in order to "influence the characteristics and the size of enrolled student bodies" [@RN2771, p. xiv]
- EM office controls recruiting, admissions, fin aid

__Interventions along enrollment funnel__

- Convert prospects to leads 
  - Purchase student lists
- Convert leads/inquiries to applicants
  - Email, mail, school visits
- Convert applicants to admits
  - Holistic review [e.g., @RN4146]
- Convert admits to enrollees
  - Financial aid offers [e.g., @RN2241]
::::

:::: {.column width=35%}

__Enrollment Funnel__
```{r, echo=FALSE, out.width="85%", fig.cap=""}
knitr::include_graphics("./assets/images/enroll_funnelv2.png")
```
::::

:::

### __Scholarship on Recruiting "in the Wild"__{.hide}

__Sociological ethnographies of enrollment management__

- Private Colleges: visit affluent "feeder schools" to maintain relationships with HS guidance counselors [@RN3519; @RN4407]
- For-profits: door-to-door visits to local, low-income communities of color [@cottom2017lower]

__Web-scraped the travel schedules of public research and selective private univs__

- Visit rich, mostly White public schools [@RN4758]
- Disproportionate visits to private schools [@RN5023]
- "Recruitment redlining": circuitous avoidance of Black and Latinx communities along visit paths [@RN4759] ([LA map](#){data-modal-type="image" data-modal-url="assets/images/recruitment_redlining_map.jpg"})

___But recruiting process not wholly done by colleges...___

- Third-party vendors and consultancies help colleges find and recruit students
<!-- [LA map](#){data-modal-type="image" data-modal-url="assets/images/chicago_geomarkets.png"} -->

### <span style="font-size:0.9em;">__U.S. Higher Education Market: A National Voucher System__</span> {.hide}

__Federal government__

- Takes a "steer, don't row" approach
- Subsidizes higher education by providing Title IV grants and loans to students
- Department of Education regulates Title IV institutions

__Students__

- Attend college to improve career, social, and civic outcomes
- Pay for college using: 
  - Household savings; grants and loans from federal, state, private sources

__Direct providers (Title IV institutions)__

- A mix of public, private nonprofit, for-profit providers
- Funding follows students to whichever institution they enroll in
  - A national voucher system
- Colleges have a financial incentive to enroll students, so competition for students

### __Student Lists are a Match-Making Intermediary__{.hide}

__@RN2247 on emergence of a national higher education system:__

- Information costs were primary barrier to efficient matches
  - Students want to attend best college, but don't know options or costs
  - Colleges want to enroll best students, but don't know who or where they are
- SAT causes "dramatic fall" in cost of "information about students" (p. 102)

__"Lead generation" is process of connecting merchants to potential consumers__

- "List-based" lead generation based on direct-mail model (legacy)
  - In 1972, College Board began selling lists of prospects to colleges
- "Behavioral-based" targeting (contemporary)
  - Target users of a platform while they use that platform or a partner platform

__Students who opt into College Board _Search_ are [@RN4739]:__

  - 25% more likely to enroll in a four-year university after high school
  - 31% more likely to graduate college within four years 
- Magnitude of effects largest for Black, Latinx, and first-gen college students

### __College Board Lists and Student Outcomes__{.hide}
#### _Howell, Hurwitz, Mabel et al. (2021)_

```{r, fig.align="left"}
knitr::include_graphics("./assets/images/cb-fig-1.png")
```

### __The Student List Project__{.hide}

[THIS SLIDE NEEDS LOTS OF WORK; NEED TO SHOW THE SEARCH FILTERS AS PROBLEMATIC]
Public records requests to public universities in CA, IL, MN, TX, AZ from 2016-2020

- "Order summary" specifying search filter criteria [(Sample)](https://drive.google.com/drive/u/1/folders/18CHMlUIa5OO0zRkET7eSIzYnvk4S9_U1)
- De-identified prospect-level student list [(Sample)](https://drive.google.com/drive/u/1/folders/1rDIAuHLsh3IfglSsqhNuCmYsfnbtc0tV)

__What is the relationship between using academic and geographic filters and the demographic composition of resulting student lists?__ [@RN5012]

- Analyze descriptive relationships observed in list purchases from College Board and HSLS:09 simulated relationships
- Findings: Racial (and socioeconomic) inequity results from:
  - College Board student list databases exclude non-test-takers
  - Test scores and geographic filters compound disparities in excluded prospects

MOTIVATE THIS STUDY

- Filters used by public research universities
- [Link](#){data-modal-type="image" data-modal-url="assets/images/research_univ_list_filters.png"}

# Emprical Case and Theory

### Background{.hide}
#### Geomarkets and Market Segment Model created by @RN4982

<p class="quote">Geomarkets</p>
<p style="display:none;">Boundaries [often] match formal political and educational divisions, reflecting natural channels of communication. **Each major metro area is composed of several markets, usually corresponding to the inner city, a first ring of suburbs, and an outer ring of suburbs.** In more sparsely populated areas, communities are sometimes combined in order to make the analysis meaningful. [@RN4982, p. 4].</p>

- Geomarkets carve states and metro areas into smaller geographic regions
- Meant to reproduce the recruiting territories of admissions officers
- @RN4982 has 143 Geomarkets in 3 regions; EPS has 304

**Market Segment Model**

- Predicts how demand for a college varies by Geomarket, based on characteristics of households in Geomarket
- Categorize students into 4 market segments -- local, in-state, regional, national -- based on SAT score-sending behavior
  - local: send most scores to colleges within Geomarket
  - regional: send most scores out-of-state but in-region (e.g., Midwest)
- Selective colleges should look for regional and national students
- @RN4982 create outputs that become EPS software

<!--   - Like @RN2247, except College Board creates product from it -->

### Scholarship on Quantification {.hide}
#### Discussions of correlation and homophily by @RN4975

<p class="quote">**Correlation**: extent to which variables move together</p>
<p style="display:none;">Our research has simply demonstrated what everyone has always known: communities with high levels of family income and parental education are also communities in which students have higher than average SATs and more far-reaching aspirations” [@RN4982, p. 42].</p>

- Correlations observed in training data caused by structural inequality
  - "Predicting the future on the basis of the past threatens to reify and reproduce existing inequalities" [@RN4794, p. 224]
- @RN4982 find 4 predictors of student market segment
  - Parental education, family income, SAT score, educational aspirations
  - Recommend selective colleges target high-SES Geomarkets

<p class="quote">**Homophily**: birds of a feather flock together</p>
<p style="display:none;">The layering of collegiate competition is primarily a socioeconomic layering. The hierarchical structure of collegiate competition reflects the stratified social and economic dimensions of communities from which colleges draw their students. Competition among colleges, as admissions officers have told us for so long, is a matter of keeping company with one’s peers [@RN4982, p. 72].</p>

- Homophily not the result of voluntary action; homophily programmed into algorithms that create connections
  - Match vertically ranked consumers to vertically ranked producers
- @RN4982: like colleges compete for like students (SES)
<p class="quote" style="font-size: 0.85em;">Recommend colleges target Geomarkets, schools targeted by peer colleges</p>
<p style="display:none;">On occasion, senior spokespersons for the profession worry that students outside the main [Geo]market areas remain forgotten and hence, unchallenged. Inevitably, the increasing competition for students, the expense of travel and mailings, and internal political constraints compel institutions to concentrate their efforts where they will do the most good. The result is a natural reinforcing of the basic socioeconomic patterns that gave shape in the first place to the structure of college choice [@RN4982, p. 44].</p>

<!--

### Scholarship on quantification {.hide}
#### Discussion of correlation by @RN4975[chapter 1]

**Correlation**, extent to which variables move together

- Predictive analytics in two steps: 
    1. Measure correlations in previous cases (training data)
    1. Apply training data correlations to predict future cases
- But correlations from training data are point-in-time snapshot
  - Observed correlations caused by enduring structural inequality
  - "Predicting the future on the basis of the past threatens to reify and reproduce existing inequalities" [@RN4794, p. 224]

<p class="quote">@RN4982[chapter 3]</p>
<p style="display:none;">Our research has simply demonstrated what everyone has always known: communities with high levels of family income and parental education are also communities in which students have higher than average SATs and more far-reaching aspirations” [@RN4982, p. 42].</p>

- Which variables correlated with student market segment (training data)?
  - Parental education, family income, SAT score, educational aspirations
  - Conclusion: demand for higher education is a function of social origin
- Recommendation: selective colleges target high-SES Geomarkets

### Scholarship on quantification {.hide}
#### Discussion of homophily by @RN4975[chapter 2]

**Homophily**, birds of a feather flock together

- Network science assumes homophily result of voluntary action by actors
- But homophily programmed into algorithms that create connections
- Homophily in market research/geodemography
  - Based on snapshot of existing social stratification, match vertically categorized consumers to vertically categorized producers
  - Amplifies the effect of initial stratificaiton on subsequent stratification

<p class="quote">@RN4982[chapter 4]</p>
<p style="display:none;">The layering of collegiate competition is primarily a socioeconomic layering. The hierarchical structure of collegiate competition reflects the stratified social and economic dimensions of communities from which colleges draw their students. Competition among colleges, as admissions officers have told us for so long, is a matter of keeping company with one’s peers [@RN4982, p. 72].</p>

- Conducted a social network analysis (training data):
  - 2-mode network: students (1) tie to colleges (2) by sending scores
  - Colleges *i* and *j* compete if many students send scores to both
  - Conclusion: Like colleges compete for like (social origin) students

<p class="quote">Recommendation: target Geomarkets, HS targeted by peer colleges</p>
<p style="display:none;">On occasion, senior spokespersons for the profession worry that students outside the main market areas remain forgotten and hence, unchallenged. Inevitably, the increasing competition for students, the expense of travel and mailings, and internal political constraints compel institutions to concentrate their efforts where they will do the most good. The result is a natural reinforcing of the basic socioeconomic patterns that gave shape in the first place to the structure of college choice [@RN4982, p. 44].</p>
 -->

### Research questions{.hide}

<br>

**Research interests**

- How unequal are these Geomarkets?
- Who is included/excluded if we purchase student lists and filter on Geomarkets as recommended by @RN4982

<br>

**Research questions**

1. What is the socioeconomic and racial variation between and within Geomarkets? How does this change over time?
1. What is the socioeconomic and and racial composition of included vs. excluded prospects, if "student list" purchases filter on particular Geomarkets?

# Methods

### Methods{.hide}
#### Data

**RQ1: SES and racial variation over time between, within Geomarkets**

- Census tract-level data
  - 1980, 2000 Census, 2016-2020 American Community Survey (ACS)
- Geomarket shapefiles obtained from a 2012 blog post
  - Each zip code assigned to one Geomarket
- Partial spatial assigns each Census tract to one or more Geomarkets

**RQ2: SES and racial composition of student list purchases**

- Student lists purchased by 14 public universities from 2016-2020
  - 414 student lists associated with 2.5 million prospects
    - Data: search filter criteria; and de-identified student-level data
    - Each prospect has home and HS zip code (point geometry)
- Analyze purchases that include all Geomarkets in a metro area
  - Show who included/excluded if we filter on particular Geomarkets

### Methods{.hide}
#### Research design is a quantitative case study

<div style="margin-top: 0.8em;"></div>

Cases

- Cases are metro areas associated with multiple Geomarkets (n ~ 17)
- Initial case selection based on availability of student list purchases (RQ2)
  - Prefer purchases that filter on test score, HS class, GPA, but not on student preferences (e.g., college size)
  - Ideal: purchases differ on test score range, but same on other filters
  - For some metros, don't have purchases that include all Geomarkets
- Final set of cases for manuscript main text (3 metro areas)
  - TBD; unsure about criteria
  - Conducted analyses for larger sample of metros

Analyses (RQ1 and RQ2)

- Descriptive statistics for between Geomarket variation
- Interactive maps for within Geomarket variation

# Results

```{r setup, echo=FALSE, results = 'asis', eval = FALSE}

library(tidyverse)
library(kableExtra)  # For collapse_rows() to work
library(tools)

knitr::opts_chunk$set(echo = FALSE, message = FALSE)

# Example directories
graphs_dir <- file.path(".", "results", "graphs")
tables_dir <- file.path(".", "results", "tables")

# Revised insert_figure() that does NOT print a heading
insert_figure <- function(rq, metro, graph_type, base_path = graphs_dir) {

  # We'll use 'metro' directly since it's already underscored (e.g. "los_angeles").
  fig_file  <- file.path(base_path, rq, stringr::str_c(rq, "_", metro, "_", graph_type, ".png"))
  cap_file  <- file.path(base_path, rq, stringr::str_c(rq, "_", metro, "_", graph_type, "_title.txt"))
  note_file <- file.path(base_path, rq, stringr::str_c(rq, "_", metro, "_", graph_type, "_note.txt"))

  # Print a heading/caption
  caption_text <- if (file.exists(cap_file)) readLines(cap_file, warn = FALSE) else "No Title Found"
  writeLines(paste0("#### ", caption_text))

  # Output the figure (if it exists)
  if (file.exists(fig_file)) {
    writeLines(stringr::str_c("![](", fig_file, ")\n"))
  } else {
    writeLines(paste("*(Missing figure file:", fig_file, ")*\n"))
  }

  # Print footnotes (if the .txt file exists)
  if (file.exists(note_file)) {
    notes_txt <- readLines(note_file, warn = FALSE)
    writeLines(
      stringr::str_c(
        "<div class=\"footnote\">",
        paste0(notes_txt, collapse = "</br>"),
        "</div>"
      )
    )
  }
}

insert_rq1_table <- function(metro) {
  table_df <- readRDS(file.path(tables_dir, str_c('rq1_table_', metro, '.rds')))
  table_df[is.na(table_df)] <- NA_real_

  headings <- unique(table_df$variable)
  headings_group <- rep(nrow(table_df) / length(headings), length(headings))
  names(headings_group) <- headings

  kbl(table_df %>% select(-variable), escape = F, col.names = c('', rep(c('Mean', 'SD', 'P25', 'P50', 'P75'), 3))) %>% 
    pack_rows(index = headings_group, colnum = 1, label_row_css = '') %>% 
    add_header_above(header = c(' ' = 1, '1980' = 5, '2000' = 5, '2020' = 5))
}

insert_rq2_table <- function(metro, table_type, rq_name = 'rq2') {
  table_df <- readRDS(file.path(tables_dir, str_c(rq_name, '_table_', metro, '.rds')))[[table_type]]
  table_df[is.na(table_df)] <- NA_real_
  table_df$order_ids <- as_factor(table_df$order_ids)

  orders <- str_c(if_else(str_detect(unique(table_df$order_ids), '_'), 'Orders ', 'Order '), sub('_', ' & ', unique(table_df$order_ids)), ', ', unique(table_df$test_range))
  orders_group <- table(table_df$order_ids)
  names(orders_group) <- orders

  if (table_type == 'race') {
    tables_group <- rep(nrow(table_df) / length(orders) / length(unique(table_df$table)), length(unique(table_df$table)))
    names(tables_group) <- c('Count of purchased students', 'Purchased students by geomarket', 'Purchased students by racial/ethnic group')

    return(
      kbl(
        table_df %>%
          mutate(
            across(all, \(x) format(round(x, 0), big.mark = ',')),
            across(race_known, ~ if_else(table == 'col_pct_table' & eps_codename != 'All', sprintf('%.1f', .), format(round(., 0), big.mark = ','))),
            across(white:nhpi, ~ if_else(table == 'count_table' | (table == 'col_pct_table' & eps_codename == 'All'), format(round(., 0), big.mark = ','), sprintf('%.1f', .)))
          ) %>%
          select(-table, -order_ids, -test_range),
        align = 'lrrrrrrrrr', escape = F,
        col.names = c('', 'All', 'Race known', 'White', 'Asian', 'Black', 'Hispanic', '2+ races', 'AIAN', 'NHPI')
      ) %>% 
        pack_rows(index = orders_group, colnum = 1, label_row_css = 'background-color: #eaeaea; padding: 10px 7.5px;') %>% 
        pack_rows(index = rep(tables_group, length(orders)), colnum = 1, label_row_css = '')
    )
  } else if (table_type == 'firstgen') {
    tables_group <- rep(nrow(table_df) / length(orders) / length(unique(table_df$table)), length(unique(table_df$table)))
    names(tables_group) <- c('Count of purchased students', 'Purchased students by geomarket', 'Purchased students by first-generation status')

    return(
      kbl(
        table_df %>%
          mutate(
            across(all, \(x) format(round(x, 0), big.mark = ',')),
            across(known, ~ if_else(table == 'col_pct_table' & eps_codename != 'All', sprintf('%.1f', .), format(round(., 0), big.mark = ','))),
            across(no_col:not_first, ~ if_else(table == 'count_table' | (table == 'col_pct_table' & eps_codename == 'All'), format(round(., 0), big.mark = ','), sprintf('%.1f', .)))
          ) %>%
          select(-unknown, -table, -order_ids, -test_range),
        align = 'lrrrrr', escape = F,
        col.names = c('', 'All', 'First-gen known', 'No college', 'Some college', 'Not first-gen')
      ) %>% 
        pack_rows(index = orders_group, colnum = 1, label_row_css = 'background-color: #eaeaea; padding: 10px 7.5px;') %>% 
        pack_rows(index = rep(tables_group, length(orders)), colnum = 1, label_row_css = '')
    )
  } else if (table_type == 'race_firstgen') {
    group_df <- table_df %>% group_by(order_ids, stu_race_cb) %>% summarise(n = n())
    race_group <- group_df$n
    names(race_group) <- str_replace(str_to_title(group_df$stu_race_cb), 'Or', 'or')

    return(
      kbl(
        table_df %>% 
          mutate(
            across(all, \(x) format(round(x, 0), big.mark = ',')),
            across(row_no_col:row_not_first_gen, \(x) sprintf('%.1f', x))
          ) %>% 
          select(-stu_race_cb, -order_ids, -test_range),
        align = 'lrrrr', escape = F,
        col.names = c('', 'All', 'No college', 'Some college', 'Not first-gen')
      ) %>% 
        pack_rows(index = orders_group, colnum = 1, label_row_css = 'background-color: #eaeaea; padding: 10px 7.5px;') %>% 
        pack_rows(index = race_group, colnum = 1, label_row_css = '')
    )
  }
}

# 2) Define your *underscored* metros / RQs
  metros <- c(
    "atlanta",
    "bay_area",
    "boston",
    "chicago",
    "cleveland",
    "dallas",
    "dc_maryland_virginia",
    "detroit",
    "houston",
    "long_island",
    "los_angeles",
    "miami",
    "new_york_city",
    "northern_new_jersey",
    "orange_county",
    "philadelphia",
    "san_diego"
  )

rqs <- c("rq1", "rq2")

# For RQ1, we have two plot types: "race" and "ses"
graph_types_rq1 <- c("race", "ses")

# For RQ2, we have two plot types: "race" and "firstgen"
graph_types_rq2 <- c("race", "firstgen")

# 3) Read the CSV (underscored metro column)
orders_df <- read_csv("scripts/metro_orders.csv")

# Derive the RQ2 metros from the CSV
rq2_metros <- unique(orders_df$metro)

# 4) Main loop
for (m in metros) {

  # Convert underscores -> spaces, then title case
  # But add special checks for "dc_maryland_virginia" and "bay_area"
  if (m == "dc_maryland_virginia") {
    heading_text <- "D.C., Maryland, and Virginia"
  } else if (m == "bay_area") {
    heading_text <- "Bay Area"
  } else {
    # Normal logic
    heading_text <- gsub("_", " ", m, fixed = TRUE)
    heading_text <- tools::toTitleCase(heading_text)
    heading_text <- str_c(heading_text, " area")
  }

  # Print a top-level heading for the metro area
  writeLines(str_c("## ", heading_text))

  for (rq in rqs) {

    # For clarity, parse out the numeric portion
    rq_num <- str_replace(string = rq, pattern = "rq", replacement = "")

    # ---------------------------------------------
    # RQ1 Logic
    # ---------------------------------------------
    if (rq == "rq1") {

      # Each metro has 2 plots: "race" & "ses"
      for (i in seq_along(graph_types_rq1)) {
        g <- graph_types_rq1[i]

        # The first RQ1 slide for a metro has a real heading;
        # subsequent ones are hidden slides (".hide")
        writeLines(str_c("### Research question ", rq_num, if_else(i == 1, "", " {.hide}")))
        insert_figure(rq = rq, metro = m, graph_type = g)
      }

      # Add RQ1 table
      writeLines(str_c("### Research question ", rq_num, " {.hide}"))
      writeLines(str_c("#### Racial/ethnic and socioeconomic characteristics of ", heading_text, " Geomarkets over time"))

      writeLines('<div class="table-wrapper freeze-pane">')
      writeLines(insert_rq1_table(m))
      writeLines('</div>')
      writeLines('<div class="footnote">Table Notes:<br>- Race/ethnicity categories not available in 1980 Census: Asian, non-Hispanic; Two+ races, non-Hispanic; NHPI, non-Hispanic; AIAN non-Hispanic<br>- Household income measured using 2024 CPI</div>')

      # Add RQ1 map
      writeLines(str_c("### Research question ", rq_num, " {.hide}"))
      writeLines(str_c("#### Racial/ethnic and socioeconomic characteristics of ", heading_text, " Geomarkets over time"))

      rq1_map <- str_c('./results/maps/rq1_map_', m, '.html')

      # Example special case for "los_angeles"
      if (m == "los_angeles") {
        rq1_map <- 'https://rq1-map-los-angeles.netlify.app'
      }

      writeLines(str_c('<iframe data-src="', rq1_map, '" src="about:blank" width=100% height=100% allowtransparency="true"></iframe>'))
      writeLines('<p class="fullscreen">Fullscreen</p>')
      writeLines('<div class="footnote">Figure Notes:<br>- Race/ethnicity categories not available in 1980 Census: Asian, non-Hispanic; Two+ races, non-Hispanic; NHPI, non-Hispanic; AIAN non-Hispanic<br>- Household income measured using 2024 CPI</div>')

    # ---------------------------------------------
    # RQ2 Logic (row vs. col approach + race_by_firstgen)
    # ---------------------------------------------
    } else if (rq == "rq2") {

      # Only run RQ2 code if the metro is in our RQ2 list
      if (!(m %in% rq2_metros)) next

      # Filter orders_df to just rows for this underscored metro
      df_metro <- filter(orders_df, metro == m)

      # If none, skip
      if (nrow(df_metro) == 0) next

      # Gather all the underscore-delimited order_ids
      orders_for_m <- df_metro$order_ids

      # -- HERE is the new line to remove "487927" --
      orders_for_m <- orders_for_m[ orders_for_m != "487927" ]

      get_notes <- function(missing_txt = '') {
        str_replace(
          str_c('<div class="footnote">', paste0(readLines(file.path(graphs_dir, rq, str_c('rq2_', m, '_race_row_plot_note.txt')), warn = FALSE), collapse = "</br>"), '</div>'),
          'Excludes students with missing values for race\\. ', missing_txt
        )
      }

      # We'll do 2 group types: race & firstgen
      for (i in seq_along(graph_types_rq2)) {
        g <- graph_types_rq2[i]

        # We'll do row and col for each group type
        for (rc in c("row", "col")) {

          if (rc == "row") {
            #
            # ROW => single combined figure on one slide
            #
            writeLines(str_c("### Research question ", rq_num, if_else(i == 1, "", " {.hide}")))
            graph_type_row <- str_c(g, "_", rc, "_plot")
            insert_figure(rq = rq, metro = m, graph_type = graph_type_row)

          } else {
            #
            # COL => separate file(s) by order ID => each on its own slide
            #
            for (order_id in orders_for_m) {

              # Each order ID gets its own "###" => new Reveal.js slide
              writeLines(str_c("### Research question ", rq_num, " {.hide}"))

              graph_type_col <- str_c(g, "_", rc, "_plot_", order_id)
              insert_figure(rq = rq, metro = m, graph_type = graph_type_col)
            }

          } # end if (row vs col)

        } # end rc loop

        # Add RQ2 table
        writeLines(str_c("### Research question ", rq_num, " {.hide}"))
        writeLines(if_else(
          g == 'race',
          str_c('#### Racial/ethnic composition of purchased student profiles by Geomarket, ', heading_text),
          str_c('#### First-generation status of purchased student profiles by Geomarket, ', heading_text)
        ))

        writeLines('<div class="table-wrapper freeze-pane">')
        writeLines(insert_rq2_table(m, g))
        writeLines('</div>')
        writeLines(get_notes())

      } # end g loop

      # -------------------------------------------
      # ADDITIONAL: "race_by_firstgen" => goes to RQ2B
      # -------------------------------------------
      for (order_id in orders_for_m) {
        writeLines(str_c("### Research question ", rq_num, " {.hide}"))
        # e.g. "rq2b_los_angeles_order_448375_546954_race_by_firstgen.png"
        graph_type_rbf <- str_c("order_", order_id, "_race_by_firstgen")
        insert_figure(rq = "rq2b", metro = m, graph_type = graph_type_rbf)
      }

      # Add RQ2B table
      writeLines(str_c("### Research question ", rq_num, " {.hide}"))
      writeLines(str_c("#### First-generation status by race for ", heading_text, " Geomarkets"))

      writeLines('<div class="table-wrapper freeze-pane">')
      writeLines(insert_rq2_table(m, 'race_firstgen'))
      writeLines('</div>')
      writeLines(get_notes('Excludes students who have missing values for race or parental education. '))

      # Add RQ2 map
      writeLines(str_c("### Research question ", rq_num, " {.hide}"))
      writeLines(str_c("#### Purchased student profiles by racial/ethnic composition and first-generation status, ", heading_text))

      writeLines(str_c(
        '<iframe data-src="./results/maps/rq2_map_',
        m,
        '.html" src="about:blank" width=100% height=100% style="max-height: 75%" allowtransparency="true"></iframe>'
      ))
      writeLines('<p class="fullscreen">Fullscreen</p>')
      writeLines(get_notes())

    } # end if (rq == "rq2")

  } # end rq loop

} # end metros loop

# A special version insert_figure_aian for AI/AN that does NOT print a heading
insert_figure_aian <- function(rq, metro, graph_type, base_path = graphs_dir) {

  fig_file  <- file.path(base_path, rq, paste0(rq, "_", metro, "_", graph_type, ".png"))
  cap_file  <- file.path(base_path, rq, paste0(rq, "_", metro, "_", graph_type, "_title.txt"))
  note_file <- file.path(base_path, rq, paste0(rq, "_", metro, "_", graph_type, "_note.txt"))

  # Instead of printing a heading, let the AI/AN loop do it (#### subtitle).
  # So we skip caption_text or print it differently if you want.

  if (file.exists(fig_file)) {
    writeLines(paste0("![](", fig_file, ")\n"))
  } else {
    writeLines(paste("*(Missing figure file:", fig_file, ")*\n"))
  }

  # Print footnotes (if the .txt file exists)
  if (file.exists(note_file)) {
    notes_txt <- readLines(note_file, warn = FALSE)
    writeLines(
      paste0(
        "<div class=\"footnote\">",
        paste0(notes_txt, collapse = "</br>"),
        "</div>"
      )
    )
  }
}

############################################################################
# 5) AI/AN results loop
############################################################################

# (Level-1) "AI/AN students"
writeLines("# AI/AN students\n")

# The five metros (with 'nyny' last) and the single Native American order
native_metros <- c("chicago", "dallas", "houston", "los_angeles", "nyny")
native_order  <- "487927"

# We want exactly 3 slides per metro, in this order:
var_rc_pairs <- list(
  c("firstgen", "row"),
  c("firstgen", "col"),
  c("race",     "row")
)

for (m in native_metros) {

  # (Level-2) heading for each metro (not a slide at slide-level=3)
  heading_text <- if (m == "nyny") {
    "New York City"
  } else {
    m_no_underscores <- gsub("_", " ", m, fixed = TRUE)
    m_title          <- tools::toTitleCase(m_no_underscores)
    paste0(m_title, " area")
  }
  writeLines(paste0("## ", heading_text, "\n"))

  # For each var/rc pair, create a new slide at level-3 (always hidden in TOC)
  for (i in seq_along(var_rc_pairs)) {

    g  <- var_rc_pairs[[i]][1]
    rc <- var_rc_pairs[[i]][2]

    # Always hidden from TOC => add "{.hide}" every time
    writeLines(paste0("### ", heading_text, " (AI/AN) {.hide}\n"))

    # (Level-4) subtitle on the same slide
    subtitle <- if (g == "firstgen") {
      "First-generation status composition of purchased AI/AN student profiles by geomarket"
    } else if (g == "race") {
      "Racial composition of purchased AI/AN student profiles by geomarket"
    } else {
      "Other composition of purchased AI/AN student profiles"
    }
    writeLines(paste0("#### ", subtitle, "\n"))

    # Build figure name and call your "no-heading" figure function
    graph_type <- paste0(g, "_", rc, "_plot_", native_order)
    insert_figure_aian(
      rq         = "rq2",
      metro      = m,
      graph_type = graph_type
    )

    # Add RQ2 table
    if ((g == 'firstgen' && rc == 'col') | (g == 'race' && rc == 'row')) {
      writeLines(paste0("### ", heading_text, " (AI/AN) {.hide}\n"))
      writeLines(if_else(
        g == 'race',
        str_c('#### Racial/ethnic composition of purchased student profiles by Geomarket, ', heading_text),
        str_c('#### First-generation status of purchased student profiles by Geomarket, ', heading_text)
      ))

      writeLines('<div class="table-wrapper freeze-pane">')
      writeLines(insert_rq2_table(m, g, 'rq2_aian'))
      writeLines('</div>')
      writeLines('<div class="footnote">Figure Notes:<br>- Order number 487927 (ordered 7/19/2019): HS class 2020 OR 2021; Ethnicity/Race = American Indian or Alaska Native; SAT 1040-1600; GPA B- to A+.</div>')
    }
  }

  # Add RQ2 map
  writeLines(paste0("### ", heading_text, " (AI/AN) {.hide}\n"))
  writeLines(str_c("#### Purchased AI/AN student profiles by racial/ethnic composition and first-generation status"))

  writeLines(str_c(
    '<iframe data-src="./results/maps/rq2_aian_map_',
    m,
    '.html" src="about:blank" width=100% height=100% allowtransparency="true"></iframe>'
  ))
  writeLines('<p class="fullscreen">Fullscreen</p>')
  writeLines('<div class="footnote">Figure Notes:<br>- Order number 487927 (ordered 7/19/2019): HS class 2020 OR 2021; Ethnicity/Race = American Indian or Alaska Native; SAT 1040-1600; GPA B- to A+.</div>')
}
```

# Discussion

### Discussion{.hide}
#### Summary of results

**RQ1: SES and racial variation over time between, within Geomarkets**

- Geomarkets highly correllated with race and class in many metros
- In 1980, Black people are usually concentrated in poorest Geomarket
  - In some metros, this pattern persists (e.g., Cleveland)

**RQ2: SES and racial composition of student list purchases**

- Poor, non-white Geomarkets produce disproportionate share of Black and Hispanic students
- Poor, non-white Geomarkets produce disproportionate share of first-generation students
- Intersection of race and parental education
  - Compared to students from affluent Geomarkets, Asian and Hispanic students from low-income Geomarkets tend to be first-gen
  - Targeting non-white students in affluent Geomarkets yields mostly non-first-gen

### Discussion{.hide}
#### Contribution to scholarship

Scholarship on quantification (in education)

- Hasn't examined quantification that classifies consumers for producers
  - Structural inequality $\rightarrow$ unequal student demand for college
  - Third-party products take snapshot student demand $\rightarrow$ encourage supply-side to amplify structural inequalities observed on demand-side
- @RN5005 show effect of quantification, but not mechanisms of *how* quantification produces inequality

Scholarship on postsecondary "organizational field"

- Soc-of-ed focuses on schools (and policies) as "sorting machines"
  - But third-party vendors have structured college access for decades
- Org field is "totality of relevant actors" [@RN527, p. 148]
  - Organizational field should be focus of regulation
  - Sociology should identify high-leverage actors/products in org field

# Thank you!{.align-center}

<div><img class="qr" src="./assets/images/qr_code.png" /></div>

# References

::: {#refs}
:::