### 2025-10-27 scripts/geomarket_cbsa.Rmd
add EPS shapes

---
title: '2020 Geomarkets + CBSA'
output:
  html_document:
    toc: true
    toc_float:
      collapsed: false
    df_print: paged
---

```{r setup, include = F}
options(width = 10000)
knitr::opts_chunk$set(warning = F, message = F, echo = F, comment = '')

library(tidyverse)
library(leaflet)

scripts_dir <- '.'

eps_data_dir <- file.path('..', 'data', 'eps_market')
```

```{r}
# EPS shapes (2020)
load(file.path(eps_data_dir, 'eps_shapes_2020.RData'))

# County shapes (2020)
load(file.path(eps_data_dir, 'counties_2020.RData'))

# CBSA shapes (2020)
load(file.path(eps_data_dir, 'cbsa_2020.RData'))
```

```{r}
plot_map <- function(eps_codes, palette = c(RColorBrewer::brewer.pal(11, 'Spectral'), 'black', 'white')) {
  state <- unique(str_match(eps_codes, '^[A-Z]{2}')[,1])

  if (length(state) > 1) print(state)

  cbsas <- cbsa_2020 %>% filter(str_detect(NAME, str_c(', ', state)) | str_detect(NAME, str_c('-', state)))
  metros <- cbsas %>% filter(LSAD == 'M1')
  micros <- cbsas %>% filter(LSAD == 'M2')

  m <- leaflet() %>% 
    addProviderTiles(providers$CartoDB.Positron) %>% 
    addPolygons(data = counties_2020 %>% filter(STUSPS %in% state) %>% sf::st_transform(crs = 4326), label = ~NAMELSAD, color = 'gray', weight = 1, group = 'Counties')

  if (nrow(metros) > 0) {
    m <- m %>%
      addPolygons(data = metros %>% sf::st_transform(crs = 4326), label = ~NAMELSAD, color = 'gray', weight = 1, group = 'CBSA (Metro Area)')
  }

  if (nrow(micros) > 0) {
    m <- m %>%
      addPolygons(data = micros %>% sf::st_transform(crs = 4326), label = ~NAMELSAD, color = 'gray', weight = 1, group = 'CBSA (Micro Area)')
  }

  bbox <- sf::st_bbox(eps_2020 %>% filter(eps %in% eps_codes)) %>% as.vector()

  for (i in seq_along(eps_codes)) {
    m <- m %>% 
      addPolygons(data = eps_2020 %>% filter(eps == eps_codes[[i]]), label = ~paste0('<b>', eps, '</b>') %>% lapply(htmltools::HTML), color = if_else(i %in% 5:7, '#c4c4c4', palette[[i]]), fillColor = palette[[i]], weight = 1, group = 'EPS from 2020 tracts')
  }

  m <- m %>% 
    addPolylines(data = eps_2020 %>% filter(eps %in% eps_codes), color = 'black', opacity = 1, weight = 2, group = 'EPS from 2020 tracts') %>% 
    addLayersControl(
      position = 'bottomleft',
      overlayGroups = c('EPS from 2020 tracts', 'CBSA (Metro Area)', 'CBSA (Micro Area)', 'Counties'),
      options = layersControlOptions(collapsed = F)
    ) %>% 
    fitBounds(bbox[1], bbox[2], bbox[3], bbox[4]) %>% 
    hideGroup('CBSA (Metro Area)') %>% 
    hideGroup('CBSA (Micro Area)') %>% 
    hideGroup('Counties')

  m
}
```

# All CBSAs

- **Metro Areas**: Pink
- **Micro Areas**: Gray

```{r}
leaflet() %>% addProviderTiles(providers$CartoDB.Positron) %>%
  addPolygons(data = cbsa_2020 %>% filter(LSAD == 'M1') %>% sf::st_transform(crs = 4326), label = ~NAMELSAD, weight = 1, opacity = 1, color = 'pink') %>%
  addPolygons(data = cbsa_2020 %>% filter(LSAD == 'M2') %>% sf::st_transform(crs = 4326), label = ~NAMELSAD, weight = 1, opacity = 1, color = '#cccccc')
```

# 1) ME, NH, VT, CT and MA (excluding Boston)

## Connecticut (CT)

- **CT 1**: New London and Windham County
- **CT 2**: New Haven and Middlesex County
- **CT 3**: Fairfield County
- **CT 4**: Waterbury and Litchfield County
- **CT 5**: Hartford and Tolland County

```{r}
plot_map(
  c('CT 1', 'CT 2', 'CT 3', 'CT 4', 'CT 5')
)
```

## Maine (ME)

- **ME 1**: Portland and Southern Maine
- **ME 2**: Augusta and Central Maine
- **ME 3**: Bangor and Norther Maine

```{r}
plot_map(
  c('ME 1', 'ME 2', 'ME 3')
)
```

## Massachusetts (MA)

- **MA 1**: Berkshire and Franklin Counties
- **MA 2**: Springfield and Hampshire County
- **MA 3**: Fitchburg and North Worcester County
- **MA11**: Worcester

```{r}
plot_map(
  c('MA 1', 'MA 2', 'MA 3', 'MA11')
)
```

## New Hampshire

- **NH 1**: Seacost
- **NH 2**: Merrimack Valley (Manchester)
- **NH 3**: Monadnock and Lake Sunapee
- **NH 4**: Lakes and White Mountain

```{r}
plot_map(
  c('NH 1', 'NH 2', 'NH 3', 'NH 4')
)
```

## Vermont

- **VT 1**: Burlington
- **VT 2**: Southern Vermont
- **VT 3**: Northern and Eastern Vermont

```{r}
plot_map(
  c('VT 1', 'VT 2', 'VT 3')
)
```

# 2) RI and Greater Boston

## Rhode Island

- **RI 1**: Providence and Northern Rhode Island
- **RI 2**: Southern Rhode Island

```{r}
plot_map(
  c('RI 1', 'RI 2')
)
```

## Greater Boston

- **MA 4**: Essex County
- **MA 5**: Cape Cod and Islands
- **MA 6**: Boston and Cambridge
- **MA 7**: Quincy and Plymouth County
- **MA 8**: Lowell, Concord and Wellesley
- **MA 9**: Norfolk and Bristol County
- **MA10**: Milton, Lexington and Waltham

```{r}
plot_map(
  c('MA 4', 'MA 5', 'MA 6', 'MA 7', 'MA 8', 'MA 9', 'MA10')
)
```

# 6) NJ (North)

## New Jersey (North)

- **NJ 7**: Union County
- **NJ 8**: Essex and Souther Passaic County
- **NJ 9**: Hudson County
- **NJ10**: Bergen County
- **NJ11**: Morris and Norther Passaic County
- **NJ12**: Sussex, Warren and Hunterdon Counties

```{r}
plot_map(
  c('NJ 7', 'NJ 8', 'NJ 9', 'NJ10', 'NJ11', 'NJ12')
)
```

# 7) NJ (South)

## New Jersey (South)

- **NJ 1**: Southern Jersey
- **NJ 2**: Camden and Burlington County
- **NJ 3**: Jersey Shore and Pinelands
- **NJ 4**: Middlesex County
- **NJ 5**: Monmouth County
- **NJ 6**: Somerset and Mercer Counties

```{r}
plot_map(
  c('NJ 1', 'NJ 2', 'NJ 3', 'NJ 4', 'NJ 5', 'NJ 6')
)
```

# 8) DE and Greater Philadeliphia

## Delaware

- **DE 1**: New Castle County
- **DE 2**: Kent and Sussex Counties

```{r}
plot_map(
  c('DE 1', 'DE 2')
)
```

## Greater Philadelphia

- **PA 1**: Bucks County
- **PA 2**: Chester County
- **PA 3**: Delaware County
- **PA 4**: Montgomery County
- **PA 5**: Philadelphia County

```{r}
plot_map(
  c('PA 1', 'PA 2', 'PA 3', 'PA 4', 'PA 5')
)
```

# 9) PA (Excluding Philadelphia)

## Pennsylvania (excluding Philadelphia)

- **PA 6**: Lehigh Valley
- **PA 7**: Northeastern Pennsylvania
- **PA 8**: North Central Pennsylvania
- **PA 9**: Northwestern Pennsylvania
- **PA10**: Southern Pennsylvania (East)
- **PA11**: Southern Pennsylvania (West)
- **PA12**: Allegheny County
- **PA13**: Unnamed

```{r}
plot_map(
  c('PA 6', 'PA 7', 'PA 8', 'PA 9', 'PA10', 'PA11', 'PA12', 'PA13')
)
```

# 10) MD

## Maryland

- **MD 1**: Western Maryland
- **MD 2**: Montgomery Metropolitan
- **MD 3**: Central Maryland excluding Baltimore
- **MD 4**: Eastern Shore
- **MD 5**: Prince Georges Metropolitan
- **MD 6**: Southern Maryland
- **MD 7**: Baltimore (Urban)

```{r}
plot_map(
  c('MD 1', 'MD 2', 'MD 3', 'MD 4', 'MD 5', 'MD 6', 'MD 7')
)
```

# 11) DC and Northern VA

## District of Columbia

- **DC 1**: District of Columbia

```{r}
plot_map(
  c('DC 1')
)
```

## Northern Virginia

- **VA 1**: Arlington and Alexandria
- **VA 2**: Fairfax County
- **VA 3**: North Central Virginia

```{r}
plot_map(
  c('VA 1', 'VA 2', 'VA 3')
)
```

# 12) KY, TN, WV and VA (excluding Northern VA)

## Kentucky

- **KY 1**: Lexington and Fayette
- **KY 2**: Louisville and Western Kentucky

```{r}
plot_map(
  c('KY 1', 'KY 2')
)
```

## Tennessee

- **TN 1**: Chattanooga
- **TN 2**: Knoxville
- **TN 3**: Memphis
- **TN 4**: Nashville and Davidson

```{r}
plot_map(
  c('TN 1', 'TN 2', 'TN 3', 'TN 4')
)
```

## West Virginia

- **WV 1**: Charleston and Huntington
- **WV 2**: Northern West Virginia

```{r}
plot_map(
  c('WV 1', 'WV 2')
)
```

## Virginia (excluding Northern VA)

- **VA 4**: Northern Neck
- **VA 5**: Central Virginia
- **VA 6**: Richmond
- **VA 7**: Southside Virginia
- **VA 8**: Tidewater
- **VA 9**: Shenandoah
- **VA10**: Southwest Virginia

```{r}
plot_map(
  c('VA 4', 'VA 5', 'VA 6', 'VA 7', 'VA 8', 'VA 9', 'VA10')
)
```

# 13) NC and SC

## North Carolina

- **NC 1**: Coastal Plains
- **NC 2**: East Central
- **NC 3**: Research Triangle
- **NC 4**: Sand Hills
- **NC 5**: North Piedmont
- **NC 6**: South Piedmont
- **NC 7**: Western North Carolina

```{r}
plot_map(
  c('NC 1', 'NC 2', 'NC 3', 'NC 4', 'NC 5', 'NC 6', 'NC 7')
)
```

## South Carolina

- **SC 1**: Pee Dee
- **SC 2**: Low Country
- **SC 3**: Mid Lands
- **SC 4**: East Piedmont
- **SC 5**: West Piedmont

```{r}
plot_map(
  c('SC 1', 'SC 2', 'SC 3', 'SC 4', 'SC 5')
)
```

# 14) Northern GA

## Northern Georgia

- **GA 1**: Cherokee, Cobb and Douglass Counties
- **GA 2**: Fulton County
- **GA 3**: DeKalb and Gwinnett Counties
- **GA 4**: Clayton, Fayette, Henry and Rockdale Counties
- **GA 5**: Northeast Georgia
- **GA 8**: Northwest Georgia

```{r}
plot_map(
  c('GA 1', 'GA 2', 'GA 3', 'GA 4', 'GA 5', 'GA 8')
)
```

# 15) LA, MS, AL, GA (South) and FL (North)

## Louisiana

- **LA 1**: Baton Rouge
- **LA 2**: New Orleans
- **LA 3**: Shreveport

```{r}
plot_map(
  c('LA 1', 'LA 2', 'LA 3')
)
```

## Mississippi

- **MS 1**: Jackson
- **MS 2**: Northern Mississippi

```{r}
plot_map(
  c('MS 1', 'MS 2')
)
```

## Alabama

- **AL 1**: Birmingham and Tuscaloosa
- **AL 2**: Huntsville and Florence
- **AL 3**: Mobile
- **AL 4**: Montgomery

```{r}
plot_map(
  c('AL 1', 'AL 2', 'AL 3', 'AL 4')
)
```

## Georgia (South)

- **GA 6**: Southeast Georgia
- **GA 7**: Southwest Georgia

```{r}
plot_map(
  c('GA 6', 'GA 7')
)
```

## Florida (North)

- **FL 1**: Panhandle
- **FL 2**: Crown
- **FL 3**: East Central

```{r}
plot_map(
  c('FL 1', 'FL 2', 'FL 3')
)
```

# 16) FL (South), AS, GU, MP, PR, UM and VI

## Florida (South)

- **FL 4**: West Central
- **FL 5**: Broward, Martin and Palm Beach Counties
- **FL 6**: Dade County
- **FL 7**: Collier, Hendry and Monroe Counties

```{r}
plot_map(
  c('FL 4', 'FL 5', 'FL 6', 'FL 7')
)
```

# 17) OH

## Ohio

- **OH 1**: Northwest Ohio
- **OH 2**: North Central Ohio
- **OH 3**: City of Cleveland (West)
- **OH 4**: City of Cleveland (East)
- **OH 5**: Cuyahoga, Geauga, and Lake Counties
- **OH 6**: Northeast Ohio
- **OH 7**: West Central Ohio
- **OH 8**: Central Ohio
- **OH 9**: Greater Cincinnati
- **OH10**: Southeast Ohio

```{r}
plot_map(
  c('OH 1', 'OH 2', 'OH 3', 'OH 4', 'OH 5', 'OH 6', 'OH 7', 'OH 8', 'OH 9', 'OH10')
)
```

# 18) MI and WI

## Michigan

- **MI 1**: Wayne County
- **MI 2**: Detroit's Northern Suburbs
- **MI 3**: Ann Arbor
- **MI 4**: Capital District
- **MI 5**: Kalamazoo and Grand Rapids
- **MI 6**: "The Thumb"
- **MI 7**: Northern Michigan

```{r}
plot_map(
  c('MI 1', 'MI 2', 'MI 3', 'MI 4', 'MI 5', 'MI 6', 'MI 7')
)
```

## Wisconsin

- **WI 1**: Madison and Janesville
- **WI 2**: Milwaukee and Racine
- **WI 3**: Northern Wisconsin

```{r}
plot_map(
  c('WI 1', 'WI 2', 'WI 3')
)
```

# 19) Greater Chicago

## Greater Chicago

- **IL 7**: Chain of Lakes
- **IL 8**: Northwest Suburbs
- **IL 9**: North Shore
- **IL10**: Evanston and Skokie
- **IL11**: City of Chicago
- **IL12**: Western Suburbs
- **IL13**: South and Southwest Suburbs

```{r}
plot_map(
  c('IL 7', 'IL 8', 'IL 9', 'IL10', 'IL11', 'IL12', 'IL13')
)
```

# 20) IL and IN (excluding Chicago)

## Illinois

- **IL 1**: Rockford
- **IL 2**: Quad Cities
- **IL 3**: Peoria
- **IL 4**: Springfield
- **IL 5**: Decatur and Champaign
- **IL 6**: Southern Illinois

```{r}
plot_map(
  c('IL 1', 'IL 2', 'IL 3', 'IL 4', 'IL 5', 'IL 6')
)
```

## Indiana

- **IN 1**: "The Region"
- **IN 2**: Northwest Indiana
- **IN 3**: South Bend and Elkhart
- **IN 4**: Northeast Indiana
- **IN 5**: West Central Indiana
- **IN 6**: East Central Indiana
- **IN 7**: Greater Indianapolis
- **IN 8**: West Indiana
- **IN 9**: South Central Indiana
- **IN10**: East Indiana
- **IN11**: Southwest Indiana
- **IN12**: Southeast Indiana

```{r}
plot_map(
  c('IN 1', 'IN 2', 'IN 3', 'IN 4', 'IN 5', 'IN 6', 'IN 7', 'IN 8', 'IN 9', 'IN10', 'IN11', 'IN12')
)
```

# 21) ND, SD, MN, IA, NE, KS, MO, AR and OK

## North Dakota

- **ND 1**: Fargo and Eastern North Dakota
- **ND 2**: Western North Dakota

```{r}
plot_map(
  c('ND 1', 'ND 2')
)
```

## South Dakota

- **SD 1**: Souix Falls and Eastern South Dakota
- **SD 2**: Western South Dakota

```{r}
plot_map(
  c('SD 1', 'SD 2')
)
```

## Minnesota

- **MN 1**: Twin Cities
- **MN 2**: Northern Minnesota

```{r}
plot_map(
  c('MN 1', 'MN 2')
)
```

## Iowa

- **IA 1**: Cedar Rapids and Eastern Iowa
- **IA 2**: Des Moines and Western Iowa

```{r}
plot_map(
  c('IA 1', 'IA 2')
)
```

## Nebraska

- **NE 1**: Lincoln
- **NE 2**: Omaha
- **NE 3**: Western Nebraska

```{r}
plot_map(
  c('NE 1', 'NE 2', 'NE 3')
)
```

## Kansas

- **KS 1**: Kansas City and Topeka
- **KS 2**: Wichita and Western Kansas

```{r}
plot_map(
  c('KS 1', 'KS 2')
)
```

## Missouri

- **MO 1**: Kansas City and St. Joseph
- **MO 2**: St. Louis and Eastern Missouri
- **MO 3**: Springfield and Southern Missouri

```{r}
plot_map(
  c('MO 1', 'MO 2', 'MO 3')
)
```

## Arkansas

- **AR 1**: Little Rock
- **AR 2**: Northern Arkansas

```{r}
plot_map(
  c('AR 1', 'AR 2')
)
```

## Oklahoma

- **OK 1**: Oklahoma City and Western Oklahoma
- **OK 2**: Tulsa and Western Oklahoma

```{r}
plot_map(
  c('OK 1', 'OK 2')
)
```

# 22) TX (excluding Dallas-Ft. Worth and Houston)

## Texas (excluding D-FW and Houston)

- **TX 1**: Amarillo, Panhandle and South Plains
- **TX 2**: El Paso
- **TX 3**: Midland, Odessa, and Trans Pecos
- **TX 4**: Abilene and San Angelo
- **TX 5**: Red River Area
- **TX 6**: Austin and Central Texas
- **TX 7**: Waco, Temple and Killeen
- **TX 8**: East Texas
- **TX 9**: Beaumont and Port Arthur
- **TX10**: Central Gulf Coast, Wharton County and Victoria County
- **TX11**: South Texas Valley
- **TX13**: Del Rio, Uvalde County, and Bexar County Area
- **TX14**: City of San Antonio

```{r}
plot_map(
  c('TX 1', 'TX 2', 'TX 3', 'TX 4', 'TX 5', 'TX 6', 'TX 7', 'TX 8', 'TX 9', 'TX10', 'TX11', 'TX13', 'TX14')
)
```

# 23) Greater Dallas-Ft.Worth

## Greater Dallas-Ft.Worth

- **TX19**: City of Dallas
- **TX20**: City of Ft. Worth
- **TX21**: Irving, Arlington and Grand Prairie
- **TX22**: Dallas County excluding City of Dallas
- **TX23**: Collin and Rockwall Counties
- **TX24**: Counties West of Dallas/Ft. Worth Metroplex

```{r}
plot_map(
  c('TX19', 'TX20', 'TX21', 'TX22', 'TX23', 'TX24')
)
```

# 24) Greater Houston

## Greater Houston

- **TX12**: Brazos and Trinity Valley
- **TX15**: Northwest Houston and Conroe School District
- **TX16**: Southwest Houston Metro Area
- **TX17**: City of Houston (East)
- **TX18**: Galveston and East Harris Counties

```{r}
plot_map(
  c('TX12', 'TX15', 'TX16', 'TX17', 'TX18')
)
```

# 25) AZ, NM, CO & UT

## Arizona

- **AZ 1**: Phoenix
- **AZ 2**: Tuscon
- **AZ 3**: Northern Arizona

```{r}
plot_map(
  c('AZ 1', 'AZ 2', 'AZ 3')
)
```

## New Mexico

- **NM 1**: Albuquerque and Northern New Mexico
- **NM 2**: Southern New Mexico

```{r}
plot_map(
  c('NM 1', 'NM 2')
)
```

## Colorado

- **CO 1**: Colorado Springs and Southeastern Colorado
- **CO 2**: Metro and Northeastern Colorado
- **CO 3**: Mountain Western Colorado

```{r}
plot_map(
  c('CO 1', 'CO 2', 'CO 3')
)
```

## Utah

- **UT 1**: Salt Lake City, Ogden & Provo
- **UT 2**: Southern Utah

```{r}
plot_map(
  c('UT 1', 'UT 2')
)
```

# 26) HI and Southern CA (excluding Los Angeles)

## Southern California

- **CA12**: Central Coast
- **CA13**: Santa Barbara and West Ventura Counties
- **CA27**: Riverside, San Bernardino and Ontario
- **CA28**: South Orange County
- **CA29**: North San Diego County excluding San Diego
- **CA30**: South San Diego County excluding San Diego
- **CA31**: City of San Diego
- **CA32**: Central Valley North
- **CA33**: Central Valley South
- **CA34**: Greater Imperial Valley

```{r}
plot_map(
  c('CA12', 'CA13', 'CA27', 'CA28', 'CA29', 'CA30', 'CA31', 'CA32', 'CA33', 'CA34')
)
```

# 27) Greater Los Angeles

## Greater Los Angeles

- **CA14**: San Fernando Valley (West)
- **CA15**: San Fernando Valley (East)
- **CA16**: Glendale and Pasadena
- **CA17**: West Los Angeles and West Beach
- **CA18**: Hollywood and Wilshire
- **CA19**: East Los Angeles
- **CA20**: South Bay
- **CA21**: South and Central Los Angeles
- **CA22**: Long Beach
- **CA23**: Covina and West Covina
- **CA24**: Whittier and North Orange County
- **CA25**: Anaheim
- **CA26**: Santa Ana

```{r}
plot_map(
  c('CA14', 'CA15', 'CA16', 'CA17', 'CA18', 'CA19', 'CA20', 'CA21', 'CA22', 'CA23', 'CA24', 'CA25', 'CA26')
)
```

# 28) NV and Northern CA

## Nevada

- **NV 1**: Las Vegas
- **NV 2**: Reno

```{r}
plot_map(
  c('NV 1', 'NV 2')
)
```

## Northern California

- **CA 1**: Far Northern California
- **CA 2**: Valley of the Moon
- **CA 3**: Sacramento County
- **CA 4**: Marin County
- **CA 5**: San Francisco County
- **CA 6**: Contra Costa County
- **CA 7**: City of Oakland
- **CA 8**: Alameda County excluding Oakland
- **CA 9**: San Mateo County
- **CA10**: City of San Jose
- **CA11**: Santa Clara County excluding San Jose

```{r}
plot_map(
  c('CA 1', 'CA 2', 'CA 3', 'CA 4', 'CA 5', 'CA 6', 'CA 7', 'CA 8', 'CA 9', 'CA10', 'CA11')
)
```

# 29) WA, OR, ID, MT, WY and AK

## Washington

- **WA 1**: Greater Seattle
- **WA 2**: South Sound
- **WA 3**: Greater Spokane
- **WA 4**: Greater Washington (East)
- **WA 5**: Greater Washington (West)
- **WA 6**: Bellingham

```{r}
plot_map(
  c('WA 1', 'WA 2', 'WA 3', 'WA 4', 'WA 5', 'WA 6')
)
```

## Oregon

- **OR 1**: Greater Portland (West)
- **OR 2**: Greater Portland (East)
- **OR 3**: Northern Valley
- **OR 4**: Southern Valley
- **OR 5**: Southwest Valley
- **OR 6**: East Oregon

```{r}
plot_map(
  c('OR 1', 'OR 2', 'OR 3', 'OR 4', 'OR 5', 'OR 6')
)
```

## Idaho

- **ID 1**: Boise City
- **ID 2**: Northern Idaho

```{r}
plot_map(
  c('ID 1', 'ID 2')
)
```

## Montana

- **MT 1**: Billings and Eastern Montana
- **MT 2**: Western Montana

```{r}
plot_map(
  c('MT 1', 'MT 2')
)
```

## Wyoming

- **WY 1**: Casper and Chayenne
- **WY 2**: Western Wyoming

```{r}
plot_map(
  c('WY 1', 'WY 2')
)
```

## Alaska

- **AK 1**: Anchorage, Kenai, and Mat-su District
- **AK 2**: Greater Alaska

```{r}
plot_map(
  c('AK 1', 'AK 2')
)
```